---
title: "Texas HB2 Article 4: Special Education (Sections 4.27-4.40) - Incentives and Likely Results"
author: "Dan Swart and Claude Sonnet 4.0"
description: "Analysis covers Article 4 (Special Education) only - Sections 4.27 through 4.40. Analysis focuses on parent-directed services programs, provider autonomy protections, IEP committee duties, commissioner determination finality, and dyslexia teacher training grants. Document spans pages 37-40 of legislative text."
format:
    html:
      toc: true
      toc-float: true
      page-layout: article
      embed-resources: true
      code-fold: true
      code-summary: "Show the code"
      code-tools: true
      code-overflow: wrap
      code-block-bg: "#FAEBD7"
      code-block-border-left: "#31BAE9"
      code-link: true
      fig-width: 8
      fig-height: 6
      fig-align: center
      html-math-method: katex
      css: swart-20250327.css
    pdf:
      toc: true
      number-sections: true
      colorlinks: true
      papersize: letter
      geometry:
        - margin=1in
      fig-width: 12
      fig-height: 10
      fig-pos: 'H'
    typst:
      toc: true
      fig-width: 12
      fig-height: 10
      keep-tex: true
      prefer-html: true
---

```{r}
#| label: setup-article4-sections27-40
#| include: false
#| echo: false
#| message: false
#| warning: false

library(tidyverse)  # For dplyr, ggplot, and friends
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(DiagrammeR) # For complete control of the layout
library(knitr)      # For controlling rendering
library(kableExtra) # For tables summarizing results
library(DT)         # For rendering content that kableExtra cannot
library(dplyr)      # For data wrangling
```
<br>
<br>

## Opinionated Executive Summary and Counter-Intuitive Implications
The legislative code analyzed in Article 4's parent-directed services and dyslexia training framework reveals a systematic dismantling of local educational judgment disguised as "parental choice" and "specialized support." All code sections are evaluated based on my own "Table of Codes for Destructive Results We Will Actually Get" (see Table below).

Disturbing is that 100% of incentives contain Code C (failure by central planning) and Code G (reduced local control), indicating complete state takeover of decisions that should belong to educators and families who actually know the children.


#### The Parent Choice-Family Dependency Paradox:

Parent-Directed Services Program promises unprecedented parental autonomy through state-funded accounts for supplemental services, flexible provider selection, and religious liberty protections, yet my coding reveals these policies will create failure by central planning (Code C in 100% of grants) while systematically replacing family independence with bureaucratic oversight. Like requiring parents to get government permission before choosing their children's tutors, these programs transform natural family decision-making into compliance theater where parents must prove their need to access their own assigned accounts.


#### The Provider Autonomy-State Approval Contradiction:

Provider and Vendor Criteria appear to protect religious liberty and institutional independence through explicit non-interference language and burden-of-proof requirements, yet the analysis shows systematic tampering (Code I in 100% of provider policies) that will destabilize existing successful relationships between families and service providers. 

Parents who have spent years building trust with specific therapists or tutors will face state-approved provider networks instead of choosing based on their children's actual needs, creating the equivalent of requiring government licensing before parents can hire babysitters.


#### The Professional Development-Training Industry Multiplication:

The Dyslexia Teacher Training Grant Program promises evidence-based capacity building through high-quality training and credentialing programs with $350 million in potential funding, yet shows the highest concentration of negative codes with financial perverse incentives (Code A and B in 75% of training incentives) plus complete bureaucratic burden creation (Code L in 100% of cases). 

Instead of improving reading instruction, this creates an elaborate professional development machinery that rewards compliance documentation over direct student support, ensuring the money never reaches classrooms while enriching the training industry that profits from teacher certification requirements.


#### The Evidence-Based Support-Standardized Intervention Inversion:

Grant Program Implementation appears to target resources toward districts lacking dyslexia capacity through stakeholder panels and evidence-based design requirements, yet Code F (individual vs system focus problems) combined with failure by central planning (Code C) will push students into standardized intervention pathways rather than truly individualized approaches. 

The ultimate irony: policies claiming to build local capacity actually eliminate the local flexibility necessary for reading interventions that match each child's specific learning profile and community resources.


#### The Religious Liberty-Regulatory Expansion Escalation:

Participant Autonomy Protections promise comprehensive religious liberty safeguards through explicit non-interference language, burden-of-proof requirements, and least-restrictive-means standards, yet they systematically create regulatory infrastructure (Code L in 100% of autonomy policies) that benefits compliance administrators more than families.

Instead of protecting genuine religious freedom, these policies create a special education legal framework where families must navigate complex regulatory systems to exercise rights they should have naturally, like requiring constitutional law degrees to attend church.


#### The Final Determination-Appeal Elimination Amplification:

Final Determination Authority promises efficient program administration through commissioner authority and elimination of appeals processes, yet Code I (system tampering) appears in 100% of determination policies, indicating these programs will destabilize existing functional relationships between families and districts. 

Parents who have spent years working collaboratively with schools around their children's needs will be forced into state-administered systems with no recourse for disagreements, creating the educational equivalent of requiring federal approval for every family decision with no appeals process.


#### The Account Access-Bureaucratic Navigation Multiplication:

IEP Committee Information Requirements promise comprehensive parent information about supplemental services and account access instructions, yet the analysis shows Code G (reduced local control) in 100% of information incentives, meaning support will be achieved by eliminating the local decision-making authority that actually creates effective special education outcomes. 

Like requiring parents to get driving directions from government agencies before visiting their children's schools, these policies assume families are better served by regulatory guidance than by direct relationships with knowledgeable professionals.


#### Summary

These contradictions reveal a parent-directed services system that has abandoned the development of family autonomy, provider relationships, and individualized programming in favor of centralized account management, procedural documentation, and bureaucratic coordination mechanisms.

Like trying to improve family dinners by requiring state approval of every recipe and dining room conversation, these policies assume students with disabilities are better served by regulatory oversight than by knowledgeable families working directly with trusted providers.

They systematically undermine the local decision-making authority, stable service relationships, and individualized programming flexibility that actually create effective special education outcomes, while creating a parent-directed services industrial complex that costs hundreds of millions in state funding while producing administrators and consultants trained to navigate bureaucratic systems rather than to build relationships with students and families around individual needs and strengths.

<br>
<br>

## Detailed Analysis

### Main Table

```{r}
#| label: main-analysis-table-article4-sections27-40

# Create analysis with coded negative results for Article 4 Sections 4.27-4.40
article4_likely_results <- data.frame(
  Category = c(
    "Parent-Directed Services Program Implementation",
    "Parent-Directed Services Program Implementation",
    "Parent-Directed Services Program Implementation",
    "Parent-Directed Services Application System",
    "Parent-Directed Services Provider Criteria",
    "Parent-Directed Services Provider Criteria",
    "Parent-Directed Services Vendor Criteria",
    "Parent-Directed Services Vendor Criteria",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "Parent-Directed Services Program Participant Autonomy",
    "IEP Committee Information Requirements",
    "IEP Committee Information Requirements",
    "Parent-Directed Services Final Determination Authority",
    "Parent-Directed Services Rule Authority",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program",
    "Dyslexia Teacher Training Grant Program"
  ),

  Incentive_Created = c(
    "Agency must approve each student meeting eligibility requirements and assign account for supplemental service purchases",
    "If applications exceed available funding, agency must award grants in order received and place remaining students on waitlist",
    "Agency must maintain online user-friendly application system for parents to apply for grants",
    "Agency must maintain online user-friendly application system for parents to apply for grants",
    "Agency must establish criteria for agency approval of providers for professional services as supplemental services",
    "Agency must provide procedures for providers to apply to become agency-approved providers",
    "Agency may establish criteria for agency approval of vendors for instructional materials categories",
    "If criteria established, agency must provide procedures for vendors to apply for agency approval",
    "Providers and vendors receiving program money are not federal financial assistance recipients based on that money alone",
    "Rules cannot consider participant or provider actions as state government agent actions",
    "Rules cannot limit provider ability to determine methods or exercise religious/institutional values",
    "Rules cannot limit participant ability to determine educational content or exercise religious values",
    "Rules cannot obligate providers or participants to act contrary to religious or institutional values",
    "Rules cannot impose regulations beyond those necessary to enforce program requirements",
    "Rules cannot require creed, practice, policy, curriculum, standard, or assessment modifications as program condition",
    "In legal challenges, state has burden of proof to show rule necessity, non-violation, and no undue burden",
    "Rules must be least restrictive means of accomplishing program purpose while recognizing provider independence",
    "ARD committees must provide parents information on supplemental services and instructional materials available",
    "ARD committees must provide instructions on accessing assigned account for supplemental service purchases",
    "Determination of commissioner under parent-directed services subchapter is final and may not be appealed",
    "Commissioner must adopt rules as necessary to administer supplemental services and instructional materials program",
    "Commissioner must establish grant program for districts to increase local capacity for dyslexia services",
    "Grants must incorporate evidence-based design and increase capacity through high-quality training or credentialing programs",
    "External stakeholder panel must assist in grant selection including parents of students with disabilities",
    "Grant recipients may accept gifts, grants, and donations to implement and administer programs",
    "Recipients cannot require financial contribution from parents to implement and administer grants",
    "Regional education service centers may administer grants awarded under dyslexia training program",
    "Commissioner may adopt rules as necessary to implement dyslexia teacher training grant program",
    "Commissioner and grant recipients may accept gifts, grants, and donations from public or private sources"
  ),

  Intended_Consequence = c(
    "Streamline approval process while maintaining quality standards for student eligibility and account assignment",
    "Ensure equitable access to programs when demand exceeds available funding through fair allocation processes",
    "Provide accessible application process that accommodates parents of varying technological abilities",
    "Make parent-directed services program easily accessible through user-friendly online application and account systems",
    "Establish professional standards and approval processes for service providers to ensure quality",
    "Create clear pathway for qualified providers to participate while maintaining program integrity",
    "Establish quality standards for instructional material vendors when deemed necessary by commissioner",
    "Provide vendor approval processes that ensure appropriate materials while maintaining market flexibility",
    "Clarify that program participation does not create federal funding compliance obligations",
    "Prevent government overreach by establishing clear boundaries for regulatory authority over private participants",
    "Preserve provider educational autonomy and religious liberty within program participation",
    "Protect participant rights to educational self-determination and religious value consistency",
    "Ensure providers can operate consistently with religious beliefs without program participation conflicts",
    "Limit government regulatory authority to only essential program operational requirements",
    "Protect institutional autonomy by preventing forced organizational or philosophical changes",
    "Establish clear legal burden of proof requirements for government regulatory actions affecting program participants",
    "Ensure government regulations are minimally intrusive while accomplishing legitimate program purposes",
    "Ensure parents receive comprehensive information about supplemental service options available through program",
    "Provide clear instructions for accessing and utilizing assigned accounts for supplemental service purchases",
    "Eliminate appeals process to provide efficient program administration and clear regulatory authority",
    "Provide regulatory framework and implementation authority for effective program administration",
    "Build local capacity to serve students with dyslexia through evidence-based training and credentialing programs",
    "Improve dyslexia services by providing high-quality professional development and specialized training opportunities",
    "Ensure grant selection process includes input from stakeholders who understand needs of students with disabilities",
    "Encourage community support and additional resources for dyslexia training programs through private contributions",
    "Ensure program accessibility by preventing financial barriers for parents seeking dyslexia services",
    "Allow regional coordination and administration of dyslexia training grants for local efficiency",
    "Provide regulatory authority for effective implementation and administration of dyslexia training programs",
    "Enable comprehensive funding and support for dyslexia training programs through diverse funding sources"
  ),

  Likely_Result = c(
    "C, G, I, L", # Student approval: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Funding allocation: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Application maintenance: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L, M", # Application system: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Provider criteria: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Provider procedures: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Vendor criteria: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Vendor procedures: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Federal assistance: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Action restrictions: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Provider limits: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Participant limits: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Value obligations: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Regulation restrictions: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Modification requirements: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Legal burden: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Restrictive means: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # ARD information: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Account instructions: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Final determination: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Rule authority: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, C, G, I, K, L, M", # Grant program: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, tax burden, bureaucracy, waste resources
    "A, B, C, G, I, L", # Evidence-based design: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Stakeholder panels: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, C, G, I, L", # Gift acceptance: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Financial restrictions: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Regional administration: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Implementation rules: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, C, G, I, L" # Donation acceptance: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, bureaucracy
  )
)

# Create the main analysis table with coded results
article4_main_table <- datatable(
  article4_likely_results,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '150px', targets = 0),   # Category  
      list(width = '300px', targets = 1),   # Incentive Created
      list(width = '250px', targets = 2),   # Intended Result
      list(width = '150px', targets = 3)    # Coded Results
    ),
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover',
  colnames = c('Category', 'Incentive Created', 'Intended Result', 'Codes for Destructive Results We Will Actually Get')
) %>%
  formatStyle(
    'Category',
    backgroundColor = styleEqual(
      c("Parent-Directed Services Program Implementation", "Parent-Directed Services Application System", "Parent-Directed Services Provider Criteria", 
        "Parent-Directed Services Vendor Criteria", "Parent-Directed Services Program Participant Autonomy", "IEP Committee Information Requirements",
        "Parent-Directed Services Final Determination Authority", "Parent-Directed Services Rule Authority", "Dyslexia Teacher Training Grant Program"),
      c('#FFE8E8', '#FFD0D0', '#E8F4FD', '#E8F6FF', '#FFF2CC', '#E8F5E8', '#F0E8FF', '#FFE8F5', '#FFE0B2')
    ),
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Likely_Result',
    fontWeight = 'bold',
    fontSize = '14px',
    textAlign = 'center',
    backgroundColor = '#FFF8DC'
  )

# Display article 4 main table
article4_main_table
```

<br>
<br>
<br>

### Supplemental Tables

```{r}
#| label: supplemental-analysis-article4-sections27-40

# Create category analysis
category_analysis_a4 <- article4_likely_results %>%
  group_by(Category) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(desc(Count)) %>%
  rename(`Incentive Category` = Category,
         `Number of Incentives` = Count)

# Analyze most frequent negative result codes
code_frequency_a4 <- article4_likely_results %>%
  separate_rows(Likely_Result, sep = ", ") %>%
  count(Likely_Result, sort = TRUE) %>%
  rename(`Negative Result Code` = Likely_Result,
         `Frequency` = n) %>%
  mutate(Percentage = round(100 * Frequency / nrow(article4_likely_results), 3))

# Key policy findings focused on coded patterns
key_findings_a4 <- data.frame(
  Finding = c(
    "Most Common Negative Pattern",
    "System Tampering and Control Dominance", 
    "Bureaucratic Implementation Pattern",
    "Financial Motivation Corruption",
    "Individual vs System Focus Imbalance",
    "Most Problematic Policy Areas",
    "State Centralization Pattern",
    "Resource Waste in Program Administration"
  ),
  Description = c(
    "Code C (failure by force) appears in 100.0% of incentives, followed by Code G (reduced local control) at 100.0%",
    "System tampering affects every single incentive, indicating complete disruption of existing special education systems",
    "Administrative burdens appear in 100.0% of incentives, creating massive compliance overhead across all special education policies",
    "Code A (intrinsic to extrinsic motivation) appears in 20.7% of incentives, systematically undermining professional dedication with financial rewards",
    "No individual focus codes (F) appear in these sections, indicating focus on systemic rather than individual interventions",
    "Parent-Directed Services Program Participant Autonomy shows highest concentration with 9 separate harmful incentives in single policy area",
    "Code G (reduced local control) appears in 100.0% of incentives, representing complete state takeover of special education decisions", 
    "Code M (waste resources) appears in 10.3% of incentives, concentrated in application systems and grant program administration"
  )
)

# === DISPLAY SUPPLEMENTAL TABLES ===

# Category Analysis Header
datatable(
  data.frame(Title = "INCENTIVE CATEGORY ANALYSIS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#FF9800',
              color = 'white')

# Category Analysis Table
datatable(
  category_analysis_a4,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Number of Incentives',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(category_analysis_a4$`Number of Incentives`), '#E8F4FD')) %>%
  formatStyle('Incentive Category',
              fontWeight = 'bold',
              textAlign = 'left')

# Code Frequency Analysis Header
datatable(
  data.frame(Title = "NEGATIVE RESULT CODE FREQUENCY"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#DC143C',
              color = 'white')

# Code Frequency Table
datatable(
  code_frequency_a4,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Frequency',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(code_frequency_a4$Frequency), '#FFB6C1')) %>%
  formatStyle('Negative Result Code',
              fontWeight = 'bold',
              textAlign = 'center',
              fontSize = '16px') %>%
  formatStyle('Percentage',
              fontWeight = 'bold',
              textAlign = 'center')

# Key Findings Header
datatable(
  data.frame(Title = "KEY CODED PATTERN FINDINGS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#F44336',
              color = 'white')

# Key Findings Table
datatable(
  key_findings_a4,
  filter = 'top',
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '25%', targets = 0),
      list(width = '75%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Finding',
              fontWeight = 'bold',
              backgroundColor = '#FFE4E1',
              textAlign = 'left') %>%
  formatStyle('Description',
              textAlign = 'left')

# Create coded summary statistics
summary_statistics_a4 <- data.frame(
  Metric = c(
    "Total Incentives Analyzed",
    "Most Problematic Code", 
    "Most Frequent Code Combination",
    "Policy Categories with Code C",
    "Policy Categories with Code G",
    "Average Codes per Incentive"
  ),
  Value = c(
    nrow(article4_likely_results),
    paste0("Code C and G (both appear ", max(code_frequency_a4$Frequency), " times each)"),
    "C, G, I, L (force + control + tampering + bureaucracy)",
    sum(grepl("C", article4_likely_results$Likely_Result)),
    sum(grepl("G", article4_likely_results$Likely_Result)),
    round(mean(lengths(strsplit(article4_likely_results$Likely_Result, ", "))), 1)
  )
)

# Summary Statistics Header
datatable(
  data.frame(Title = "CODED ANALYSIS SUMMARY STATISTICS"),
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#9C27B0',
              color = 'white')

# Summary Statistics Table
datatable(
  summary_statistics_a4,
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '40%', targets = 0),
      list(width = '60%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Metric',
              fontWeight = 'bold',
              textAlign = 'left') %>%
  formatStyle('Value',
              textAlign = 'left',
              fontWeight = 'bold')
```

<br>
<br>

### Key to List of Likely Negative Results

```{r}
#| label: negative-results-coding-key-article4-sections27-40

# Create negative results table with color coding
negative_results_a4 <- data.frame(
  Code = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N"),

  Likely_Negative_Result = c(
    "Will move people from intrinsic to extrinsic motivation",
    "Will makes money the most important motivation",  
    "Will fail because improvements cannot be realized by force from the state",
    "People will become 'the problem', not the system that they work in",
    "Will decrease collaboration, pitting people against each other for rewards",
    "Will decrease opportunities and efforts to improve the system teachers work in; try to 'improve' individual teachers instead",
    "Will reduce local control",
    "Will reduce economic agility of districts when they need it most",
    "Will tamper with existing stable systems, making things worse",
    "Impedes career development when goal posts constantly moving",
    "Shifts the onus of tax increases to districts to pay for unfunded state mandates",
    "Will create more administrative and bureaucratic burdens",
    "Will cause administration and teachers to waste even more resources, while bankrupting financially responsible districts",
    "Will create long-term district obligations that are not sustainable when state funding is changed every 2 years"
  )
)

# Define bright contrasting color palette (7 colors)
color_palette_a4 <- c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF")

# Create color assignments for each row (cycling through 7 colors for 14 rows)
row_colors_a4 <- rep(color_palette_a4, length.out = nrow(negative_results_a4))

# Display the coding key table
DT::datatable(
  negative_results_a4,
  options = list(
    pageLength = 15,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    columnDefs = list(
      list(width = '40px', targets = 0),  # Narrow first column
      list(width = '90%', targets = 1)    # Wide second column
    ),
    scrollX = FALSE,
    autoWidth = FALSE
  ),
  rownames = FALSE,
  class = 'display'
) %>%
  formatStyle(
    'Code',
    textAlign = 'right',
    fontWeight = 'bold',
    backgroundColor = styleEqual(negative_results_a4$Code, row_colors_a4),
    color = 'white'
  ) %>%
  formatStyle(
    'Likely_Negative_Result',
    backgroundColor = styleEqual(negative_results_a4$Code, row_colors_a4),
    color = 'black'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    border = '2px solid #333',
    borderCollapse = 'collapse'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    maxWidth = '9in',
    tableLayout = 'fixed'
  )

```


