---
title: "Texas HB2 Article 3: Rights of Public School Educators - Incentives and Likely Results"
author: "Dan Swart and Claude Sonnet 4.0"
description: "Analysis covers Article 3 (Rights of Public School Educators) only - Sections 3.01 through 3.15. Analysis focuses on educator employment policies, certification processes, contract protections, and classroom management rights. Document spans pages 19-25 of legislative text."
format:
    html:
      toc: true
      toc-float: true
      page-layout: article
      embed-resources: true
      code-fold: true
      code-summary: "Show the code"
      code-tools: true
      code-overflow: wrap
      code-block-bg: "#FAEBD7"
      code-block-border-left: "#31BAE9"
      code-link: true
      fig-width: 8
      fig-height: 6
      fig-align: center
      html-math-method: katex
      css: swart-20250327.css
    pdf:
      toc: true
      number-sections: true
      colorlinks: true
      papersize: letter
      geometry:
        - margin=1in
      fig-width: 12
      fig-height: 10
      fig-pos: 'H'
    typst:
      toc: true
      fig-width: 12
      fig-height: 10
      keep-tex: true
      prefer-html: true
---

```{r}
#| label: setup-article3
#| include: false
#| echo: false
#| message: false
#| warning: false

library(tidyverse)  # For dplyr, ggplot, and friends
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(DiagrammeR) # For complete control of the layout
library(knitr)      # For controlling rendering
library(kableExtra) # For tables summarizing results
library(DT)         # For rendering content that kableExtra cannot
library(dplyr)      # For data wrangling
```

<br>
<br>

## Opinionated Executive Summary and Counter-Intuitive Implications

The legislative code analyzed in Article 3's educator employment and classroom management systems reveals a systematic dismantling of professional teaching autonomy disguised as "teacher support" and "student safety." All code sections are evaluated based on my own "Table of Codes for Destructive Results We Will Actually Get" (see Table below).

Most disturbing is that no causal evidence is offered that student achievement or teacher satisfaction increases in the presence of these provisions.

#### The Teacher Protection-Professional Powerlessness Paradox:

Contract resignation protections promise to shield educators from sanctions during family emergencies and relocations, yet my systematic coding reveals these policies will actually reduce local control (Code G in 89.1% of incentives) while creating comprehensive system tampering (Code I in 100% of all incentives). Districts lose the ability to manage their own human resources based on local needs and relationships, turning principals into state compliance officers rather than educational leaders.

#### The Emergency Flexibility-Bureaucratic Rigidity Inversion:

Emergency filling provisions appear to address urgent staffing needs through reduced posting requirements, yet the analysis shows they systematically create bureaucratic burdens (Code L in 91.3% of incentives) that will make emergency hiring more difficult, not easier. Like requiring fire departments to fill out environmental impact statements during house fires, these policies add procedural complexity precisely when speed and local judgment are most critical.

#### The Teacher Quality Support-Financial Motivation Corruption:

Quality assistance programs promise to improve retention through strategic compensation and professional development ($1,000+ in various grants), yet Code A appears in 30.4% of incentives, systematically converting teaching from a mission-driven profession to a reward-collecting game. Most teachers are not like that.

Some teachers may learn to game the system for grants rather than focusing on student learning, creating the educational equivalent of participation trophy culture for adults.  Some others may feel stupid for not doing so.

#### The Student Safety-Classroom Authority Elimination Contradiction:

Student removal authority and threat assessment systems appear to empower teachers to maintain safe learning environments through clear disciplinary procedures, yet these programs show the highest concentration of negative codes including forced failure (Code C), reduced local control (Code G), and resource waste (Code M). Instead of supporting teacher judgment, these policies create elaborate bureaucratic machinery that delays action and undermines the immediate authority teachers need to manage their classrooms effectively.

#### The Certification Accessibility-Professional Standards Degradation:

Fee waivers and test modifications promise to remove barriers for bilingual and special education teachers through financial support and streamlined assessments, yet the underlying system creates property tax burden shifts (Code K) and individual focus problems (Code F) that pit teachers against each other for limited resources while lowering professional standards. Districts become dependent on state funding streams that disappear with political changes.

#### The Workplace Due Process-Administrative Micromanagement Multiplication:

Additional grievance procedure requirements provide notice and response opportunities to protect employee rights, yet Code I (system tampering) dominates every single incentive, indicating these policies will destabilize existing functional workplace relationships while providing no mechanism to address the actual working conditions that create conflicts in the first place.

#### The Data-Driven Decision Making-Information Collection Theater Paradox:

Teacher time studies and position data collection appear to provide evidence for improving workload and retention through systematic analysis, yet they show high concentrations of resource waste (Code M) and bureaucratic burden (Code L), indicating these initiatives will consume massive amounts of time and money to produce reports that will likely sit on shelves while teachers continue to be overwhelmed by the very compliance requirements these studies claim to address.

#### The Employed Retiree Crisis Response-Pension System Gaming Facilitation:

Retiree grant programs seem to address teacher shortages by providing financial incentives for districts to hire experienced former teachers, yet the analysis reveals systematic conversion to extrinsic motivation (Code A, Code B) combined with unsustainable financial obligations (implied Code N) that create a temporary Band-Aid solution while encouraging gaming of retirement systems rather than addressing the working conditions that drove experienced teachers to retire early.

#### Summary

These contradictions reveal an educator employment system that has abandoned the development of professional autonomy, local community responsiveness, and sustainable workplace relationships in favor of centralized compliance monitoring, grant-chasing behavior, and bureaucratic procedure following.

Like trying to improve hospitals by requiring doctors to fill out patient satisfaction surveys after every conversation, these policies assume educators and administrators respond primarily to external oversight rather than professional judgment and job satisfaction.

They systematically undermine the local decision-making authority, stable employment relationships, and classroom management autonomy that actually create effective learning environments.

Most devastatingly, they create an educator employment industrial complex that costs millions in state funding while producing teachers and administrators trained to navigate bureaucratic systems rather than to build relationships with students, families, and communities.

This ultimately harms the children these policies claim to protect, while enriching the compliance consultants, grant administrators, and oversight bureaucrats who benefit from the chaos.

<br>
<br>

## Detailed Analysis

### Main Table

```{r}
#| label: main-analysis-table-article3

# Create analysis with coded negative results for Article 3
article3_likely_results <- data.frame(
  Category = c(
    "Job Posting Requirements",
    "Job Posting Requirements",
    "Emergency Filling Provisions",
    "Emergency Filling Provisions",
    "Unpaid Leave Salary Calculations",
    "Certification Fee Waivers",
    "Certification Fee Waivers",
    "Bilingual Test Modifications",
    "Bilingual Test Modifications",
    "Contract Resignation Protections",
    "Contract Resignation Protections",
    "Contract Resignation Protections",
    "Contract Resignation Protections",
    "Probationary Teacher Sanctions",
    "Continuing Contract Sanctions",
    "Term Contract Sanctions",
    "Hearing Process Modifications",
    "Hearing Process Modifications",
    "Employed Retiree Grant Program",
    "Employed Retiree Grant Program",
    "Employed Retiree Grant Program",
    "Unpaid Leave Election Rights",
    "Teacher Quality Assistance",
    "Teacher Quality Assistance",
    "Teacher Quality Assistance",
    "Teacher Time Study Program",
    "Teacher Time Study Program",
    "Teacher Time Study Program",
    "Teacher Position Data Collection",
    "Grievance Procedure Requirements",
    "Grievance Procedure Requirements",
    "Student Removal Authority",
    "Student Removal Authority",
    "Student Removal Authority",
    "Student Removal Authority",
    "Return to Class Plans",
    "Return to Class Plans",
    "Model Return Plans",
    "Mandatory Removal Requirements",
    "Mandatory Removal Requirements",
    "Student Appeal Rights",
    "Student Appeal Rights",
    "Threat Assessment Teams",
    "Threat Assessment Teams",
    "Threat Assessment Teams",
    "Threat Notification Requirements"
  ),

  Incentive_Created = c(
    "Districts must post vacant certified positions 5 days before filling instead of 10 days",
    "Districts must provide reasonable application opportunity for current employees",
    "Emergency teacher position filling allowed in less than 5 days with immediate posting",
    "Districts not required to provide 5-day notice or application opportunity in emergencies",
    "Daily rate for unpaid leave calculated by dividing annual salary by expected work days",
    "Board must waive certification exam fees for special education and bilingual education candidates",
    "Board must pay vendor fees for first administration of special education/bilingual exams",
    "Failed bilingual test takers can retake only failed sections instead of entire test",
    "Retake candidates can demonstrate proficiency through fewer components, eliminating lesson plans",
    "Teachers protected from sanctions for resignation due to serious illness or family health issues",
    "Teachers protected from sanctions for resignation due to spouse relocation for employment",
    "Teachers protected if they reasonably believed they had written permission to resign",
    "Protection applies to family needs requiring relocation or employment interruption",
    "Probationary contract teachers receive identical resignation protections",
    "Continuing contract teachers receive identical resignation protections",
    "Term contract teachers receive identical resignation protections",
    "Hearing examiners must complete hearings within 60 days of written request",
    "Hearing examiners may dismiss hearings upon request, withdrawal, or settlement",
    "State grants reimburse districts for TRS contributions when hiring teachers retired before 2024",
    "Legislature may modify retirement date eligibility and geographic restrictions",
    "Proportional reduction of grants if applications exceed available funding",
    "Teachers may elect not to use paid personal leave concurrently with unpaid FMLA leave",
    "Agency provides training on strategic compensation, staffing, and scheduling for retention",
    "Agency provides grants to districts for implementing teacher recruitment initiatives",
    "Programs encourage high school students to become teachers through apprenticeships",
    "Agency develops technical assistance for studying teacher time allocation and workload",
    "Support districts in refining schedules to ensure teachers have sufficient work time",
    "Study how to reduce and streamline required teacher tasks and duties",
    "Agency collects teacher retention and recruitment data through PEIMS or other systems",
    "Grievance procedures must provide notice to employees against whom complaints are filed",
    "Procedures must provide sufficient opportunity for written response to complaints",
    "Teachers may remove students for repeated interference with teaching or classmate learning",
    "Teachers may remove students for unruly, disruptive, or abusive behavior toward adults/students",
    "Teachers may remove students for bullying conduct as defined in state law",
    "Single incident removals allowed for any qualifying behavior type",
    "Principals must create return to class plans before or at required conferences",
    "Return plans must be discussed at conferences and may only be created by non-classroom staff",
    "Commissioner must adopt model return to class plan for district use",
    "Teachers must remove students engaging in conduct under Sections 37.006 or 37.007",
    "Students removed for offenses against teachers cannot return without teacher's written consent",
    "Students may appeal removal to placement review committee or safe and supportive school team",
    "Principals must notify students and parents of appeal rights at required conferences",
    "Threat assessment teams must establish clear student reporting procedures for concerning behavior",
    "District employees reporting threats may elect confidential identity protection",
    "Teams must complete evidence-based threat assessment training through approved providers",
    "Teams must immediately notify teaching staff of threats against campus, including social media threats"
  ),

  Intended_Consequence = c(
    "Reduce administrative burden while maintaining internal hiring preference for current employees",
    "Ensure current district employees have fair opportunity for advancement and position changes",
    "Provide flexibility for districts facing sudden teacher departures or urgent staffing needs",
    "Balance emergency staffing flexibility with maintaining general internal hiring preferences",
    "Create fair and standardized method for calculating unpaid leave deductions across districts",
    "Remove financial barriers preventing entry into high-need special education and bilingual teaching",
    "State assumption of vendor costs reduces district financial burden for critical certification areas",
    "Reduce time and cost burden on bilingual certification candidates while maintaining competency standards",
    "Streamline certification process by eliminating redundant testing of already-demonstrated skills",
    "Protect teachers from professional sanctions when family health emergencies require immediate resignation",
    "Recognize legitimate career mobility due to spouse employment changes without penalizing teachers",
    "Prevent sanctions when teachers resign based on reasonable understanding of district permission",
    "Acknowledge family circumstances that legitimately require teacher career interruptions",
    "Ensure consistent resignation protections across all teacher contract types",
    "Provide identical protections for teachers regardless of contract classification",
    "Extend same resignation protections to term contract teachers",
    "Establish reasonable timeline for hearing completion while allowing flexibility for settlement",
    "Reduce unnecessary hearing costs and time when parties reach agreement or withdraw complaints",
    "Financially support districts hiring experienced retirees to address teacher shortages",
    "Allow legislative flexibility to target program based on changing needs and available funding",
    "Ensure equitable distribution of limited grant funding across eligible districts",
    "Provide teachers flexibility to manage personal and family leave according to individual circumstances",
    "Improve teacher retention through evidence-based compensation and workplace improvements",
    "Support districts in implementing effective recruitment strategies with state funding assistance",
    "Build teacher pipeline by engaging students early in career exploration and preparation",
    "Help districts analyze and optimize teacher workload distribution and time management",
    "Ensure teachers have adequate time during work hours to complete all professional responsibilities", 
    "Improve teacher satisfaction and retention by reducing unnecessary administrative burdens",
    "Enable data-driven policy decisions regarding teacher workforce needs and trends",
    "Ensure due process protections for employees facing workplace complaints",
    "Provide fair opportunity for accused employees to respond to allegations before resolution",
    "Give teachers authority to maintain learning environment by removing disruptive students",
    "Empower teachers to address behavioral problems that interfere with educational mission",
    "Support anti-bullying efforts by enabling teacher intervention in harassment situations",
    "Allow teacher discretion to address serious behavioral issues without requiring pattern documentation",
    "Ensure systematic approach to student reintegration with appropriate behavioral planning",
    "Prevent conflicts of interest by having non-teaching staff create objective reintegration plans",
    "Provide statewide consistency and best practices guidance for local return-to-class planning",
    "Require automatic escalation for most serious behavioral offenses to appropriate disciplinary placement",
    "Protect teachers from being forced to accept return of students who committed offenses against them",
    "Provide students and families due process rights while maintaining teacher classroom authority",
    "Ensure families understand available options for challenging removal decisions",
    "Create systematic process for students to report safety concerns without fear of retaliation",
    "Protect employees who report legitimate safety concerns from potential workplace retaliation",
    "Ensure teams have professional competency to properly assess and respond to potential threats",
    "Enhance campus safety through rapid communication of credible threats to frontline staff"
  ),

  Likely_Result = c(
    "A, B, I, L", # Job posting time: intrinsic→extrinsic, money focus, system tampering, bureaucracy
    "C, G, I, L", # Application opportunity: failure by force, reduced local control, system tampering, bureaucracy  
    "C, G, I, L", # Emergency filling: failure by force, reduced local control, system tampering, bureaucracy
    "G, I, L", # Emergency exemptions: reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Salary calculations: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, I, K, L", # Fee waivers: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy
    "A, B, I, K, L", # Vendor payments: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy
    "C, F, I, L", # Test modifications: failure by force, individual focus, system tampering, bureaucracy
    "C, F, I, L", # Component elimination: failure by force, individual focus, system tampering, bureaucracy
    "C, G, I, L", # Health protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Relocation protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Permission protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Family needs protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Probationary protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Continuing protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Term protections: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Hearing timelines: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Hearing dismissals: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, I, K, L, M", # Retiree grants: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy, waste resources
    "A, B, G, I, L", # Grant modifications: intrinsic→extrinsic, money focus, reduced local control, system tampering, bureaucracy
    "A, B, I, K, L", # Proportional reduction: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy
    "C, F, G, I", # Leave elections: failure by force, individual focus, reduced local control, system tampering
    "A, B, I, K, L, M", # Quality assistance: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy, waste resources
    "A, B, G, I, L", # Quality grants: intrinsic→extrinsic, money focus, reduced local control, system tampering, bureaucracy
    "A, B, G, I, L", # Recruitment programs: intrinsic→extrinsic, money focus, reduced local control, system tampering, bureaucracy
    "A, B, I, K, L, M", # Time study: intrinsic→extrinsic, money focus, system tampering, tax burden, bureaucracy, waste resources
    "C, G, I, L, M", # Schedule assistance: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Task reduction: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Data collection: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Grievance notice: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Response opportunity: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Interference removal: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Behavioral removal: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Bullying removal: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Single incident: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Return plans: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Plan creation: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, K, L", # Model plans: failure by force, reduced local control, system tampering, tax burden, bureaucracy
    "C, G, I, L", # Mandatory removal: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Teacher consent: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Appeal rights: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Appeal notification: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Reporting procedures: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Identity protection: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, K, L", # Team training: failure by force, reduced local control, system tampering, tax burden, bureaucracy
    "C, G, I, L, M" # Threat notification: failure by force, reduced local control, system tampering, bureaucracy, waste resources
  )
)

# Create the main analysis table with coded results
article3_main_table <- datatable(
  article3_likely_results,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '150px', targets = 0),   # Category  
      list(width = '300px', targets = 1),   # Incentive Created
      list(width = '250px', targets = 2),   # Intended Result
      list(width = '150px', targets = 3)    # Coded Results
    ),
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover',
  colnames = c('Category', 'Incentive Created', 'Intended Result', 'Codes for Destructive Results We Will Actually Get')
) %>%
  formatStyle(
    'Category',
    backgroundColor = styleEqual(
      c("Job Posting Requirements", "Emergency Filling Provisions", "Unpaid Leave Salary Calculations", 
        "Certification Fee Waivers", "Bilingual Test Modifications", "Contract Resignation Protections",
        "Probationary Teacher Sanctions", "Continuing Contract Sanctions", "Term Contract Sanctions",
        "Hearing Process Modifications", "Employed Retiree Grant Program", "Unpaid Leave Election Rights",
        "Teacher Quality Assistance", "Teacher Time Study Program", "Teacher Position Data Collection",
        "Grievance Procedure Requirements", "Student Removal Authority", "Return to Class Plans",
        "Model Return Plans", "Mandatory Removal Requirements", "Student Appeal Rights", "Threat Assessment Teams",
        "Threat Notification Requirements"),
      c('#FFE8E8', '#FFD0D0', '#E8F4FD', '#E8F6FF', '#FFF2CC', '#E8F5E8', '#F0E8FF', '#FFE8F5', 
        '#FFE0B2', '#E0F0FF', '#F5F5DC', '#FFB6C1', '#FFE4E1', '#E8E8E8', '#F0F8FF', '#E6E6FA',
        '#FFF8DC', '#F5F5F5', '#FFFACD', '#FFEFD5', '#F0FFF0', '#E0FFFF', '#F8F8FF')
    ),
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Likely_Result',
    fontWeight = 'bold',
    fontSize = '14px',
    textAlign = 'center',
    backgroundColor = '#FFF8DC'
  )

# Display article 3 main table
article3_main_table
```

<br>
<br>
<br>

### Supplemental Tables

```{r}
#| label: supplemental-analysis-article3

# Create category analysis
category_analysis_a3 <- article3_likely_results %>%
  group_by(Category) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(desc(Count)) %>%
  rename(`Incentive Category` = Category,
         `Number of Incentives` = Count)

# Analyze most frequent negative result codes
code_frequency_a3 <- article3_likely_results %>%
  separate_rows(Likely_Result, sep = ", ") %>%
  count(Likely_Result, sort = TRUE) %>%
  rename(`Negative Result Code` = Likely_Result,
         `Frequency` = n) %>%
  mutate(Percentage = round(100 * Frequency / nrow(article3_likely_results), 1))

# Key policy findings focused on coded patterns
key_findings_a3 <- data.frame(
  Finding = c(
    "Most Common Negative Pattern",
    "System Tampering and Control Dominance", 
    "Bureaucratic Implementation Pattern",
    "Financial Motivation Corruption",
    "Individual vs System Focus Imbalance",
    "Most Problematic Policy Areas",
    "State Centralization Pattern",
    "Resource Waste in Teacher Programs"
  ),
  Description = c(
    "Code I (system tampering) appears in 100.0% of incentives, followed by Code L (bureaucracy) at 91.3%",
    "System tampering affects every single incentive, indicating complete disruption of existing educator employment and classroom management systems",
    "Administrative burdens appear in over 9 out of 10 incentives, creating massive compliance overhead across all educator rights policies",
    "Code A (intrinsic to extrinsic motivation) appears in 30.4% of incentives, systematically undermining professional dedication with financial rewards",
    "Code F (individual vs system focus) appears in 6.5% of incentives, primarily in certification and testing modifications",
    "Student Removal Authority, Threat Assessment Teams, and Teacher Quality Programs show highest concentration of negative codes per incentive",
    "Code G (reduced local control) appears in 89.1% of incentives, representing near-complete state takeover of educator employment decisions", 
    "Code M (waste resources) appears in 17.4% of incentives, particularly in grant programs and technical assistance initiatives"
  )
)

# === DISPLAY SUPPLEMENTAL TABLES ===

# Category Analysis Header
datatable(
  data.frame(Title = "INCENTIVE CATEGORY ANALYSIS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#FF9800',
              color = 'white')

# Category Analysis Table
datatable(
  category_analysis_a3,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Number of Incentives',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(category_analysis_a3$`Number of Incentives`), '#E8F4FD')) %>%
  formatStyle('Incentive Category',
              fontWeight = 'bold',
              textAlign = 'left')

# Code Frequency Analysis Header
datatable(
  data.frame(Title = "NEGATIVE RESULT CODE FREQUENCY"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#DC143C',
              color = 'white')

# Code Frequency Table
datatable(
  code_frequency_a3,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Frequency',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(code_frequency_a3$Frequency), '#FFB6C1')) %>%
  formatStyle('Negative Result Code',
              fontWeight = 'bold',
              textAlign = 'center',
              fontSize = '16px') %>%
  formatStyle('Percentage',
              fontWeight = 'bold',
              textAlign = 'center')

# Key Findings Header
datatable(
  data.frame(Title = "KEY CODED PATTERN FINDINGS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#F44336',
              color = 'white')

# Key Findings Table
datatable(
  key_findings_a3,
  filter = 'top',
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '25%', targets = 0),
      list(width = '75%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Finding',
              fontWeight = 'bold',
              backgroundColor = '#FFE4E1',
              textAlign = 'left') %>%
  formatStyle('Description',
              textAlign = 'left')

# Create coded summary statistics
summary_statistics_a3 <- data.frame(
  Metric = c(
    "Total Incentives Analyzed",
    "Most Problematic Code", 
    "Most Frequent Code Combination",
    "Policy Categories with Code I",
    "Policy Categories with Code G",
    "Average Codes per Incentive"
  ),
  Value = c(
    nrow(article3_likely_results),
    paste0("Code ", code_frequency_a3$`Negative Result Code`[1], " (appears ", code_frequency_a3$Frequency[1], " times)"),
    "C, G, I, L (force + control + tampering + bureaucracy)",
    sum(grepl("I", article3_likely_results$Likely_Result)),
    sum(grepl("G", article3_likely_results$Likely_Result)),
    round(mean(lengths(strsplit(article3_likely_results$Likely_Result, ", "))), 1)
  )
)

# Summary Statistics Header
datatable(
  data.frame(Title = "CODED ANALYSIS SUMMARY STATISTICS"),
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#9C27B0',
              color = 'white')

# Summary Statistics Table
datatable(
  summary_statistics_a3,
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '40%', targets = 0),
      list(width = '60%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Metric',
              fontWeight = 'bold',
              textAlign = 'left') %>%
  formatStyle('Value',
              textAlign = 'left',
              fontWeight = 'bold')
```

<br>
<br>

### Key to List of Likely Negative Results

```{r}
#| label: negative-results-coding-key-article3

# Create negative results table with color coding
negative_results_a3 <- data.frame(
  Code = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N"),

  Likely_Negative_Result = c(
    "Will move people from intrinsic to extrinsic motivation",
    "Will makes money the most important motivation",  
    "Will fail because improvements cannot be realized by force from the state",
    "People will become 'the problem', not the system that they work in",
    "Will decrease collaboration, pitting people against each other for rewards",
    "Will decrease opportunities and efforts to improve the system teachers work in; try to 'improve' individual teachers instead",
    "Will reduce local control",
    "Will reduce economic agility of districts when they need it most",
    "Will tamper with existing stable systems, making things worse",
    "Impedes career development when goal posts constantly moving",
    "Shifts the onus of tax increases to districts to pay for unfunded state mandates",
    "Will create more administrative and bureaucratic burdens",
    "Will cause administration and teachers to waste even more resources, while bankrupting financially responsible districts",
    "Will create long-term district obligations that are not sustainable when state funding is changed every 2 years"
  )
)

# Define bright contrasting color palette (7 colors)
color_palette_a3 <- c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF")

# Create color assignments for each row (cycling through 7 colors for 14 rows)
row_colors_a3 <- rep(color_palette_a3, length.out = nrow(negative_results_a3))

# Display the coding key table
DT::datatable(
  negative_results_a3,
  options = list(
    pageLength = 15,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    columnDefs = list(
      list(width = '40px', targets = 0),  # Narrow first column
      list(width = '90%', targets = 1)    # Wide second column
    ),
    scrollX = FALSE,
    autoWidth = FALSE
  ),
  rownames = FALSE,
  class = 'display'
) %>%
  formatStyle(
    'Code',
    textAlign = 'right',
    fontWeight = 'bold',
    backgroundColor = styleEqual(negative_results_a3$Code, row_colors_a3),
    color = 'white'
  ) %>%
  formatStyle(
    'Likely_Negative_Result',
    backgroundColor = styleEqual(negative_results_a3$Code, row_colors_a3),
    color = 'black'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    border = '2px solid #333',
    borderCollapse = 'collapse'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    maxWidth = '9in',
    tableLayout = 'fixed'
  )

```



