---
title: "Texas HB2 Article 4: Special Education Funding and Services (Sections 4.50-4.65) - Incentives and Likely Results"
author: "Dan Swart and Claude Sonnet 4.0"
description: "Analysis covers Article 4 (Special Education) Sections 4.50 through 4.65 only. Analysis focuses on special education funding formulas, dyslexia program requirements, special transportation services, day placement programs, parent-directed services grants, regional day school programs, and various special education administrative requirements. Document spans pages 47-54 of legislative text."
format:
    html:
      toc: true
      toc-float: true
      page-layout: article
      embed-resources: true
      code-fold: true
      code-summary: "Show the code"
      code-tools: true
      code-overflow: wrap
      code-block-bg: "#FAEBD7"
      code-block-border-left: "#31BAE9"
      code-link: true
      fig-width: 8
      fig-height: 6
      fig-align: center
      html-math-method: katex
      css: swart-20250327.css
    pdf:
      toc: true
      number-sections: true
      colorlinks: true
      papersize: letter
      geometry:
        - margin=1in
      fig-width: 12
      fig-height: 10
      fig-pos: 'H'
    typst:
      toc: true
      fig-width: 12
      fig-height: 10
      keep-tex: true
      prefer-html: true
---

```{r}
#| label: setup-article4-sections50-65-revised
#| include: false
#| echo: false
#| message: false
#| warning: false

library(tidyverse)  # For dplyr, ggplot, and friends
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(DiagrammeR) # For complete control of the layout
library(knitr)      # For controlling rendering
library(kableExtra) # For tables summarizing results
library(DT)         # For rendering content that kableExtra cannot
library(dplyr)      # For data wrangling
```

<br>

### Opinionated Executive Summary:  Good Intentions, Unintended Results

The rules in Article 4, Sections 4.50-4.65 reveal several issues that show how the law will produce disappointing results.


#### Funding Confusion

The new funding system seems like it will provide more money to help special education students. However, the system is so complicated that schools will spend more time filling out forms and sorting students into categories than actually helping them learn. What was supposed to be "help" becomes an extra burden.


#### Limited Choices for Parents

Parents are promised more control over their child's education through parent-directed service grants. But the state controls the options that parents can choose from. Instead of picking what their child really needs, parents are limited to options selected by the state. It’s like being told you can pick any color car you want, but only if it’s black.


#### Even More Data Overload

Schools are required to collect tons of data about students with dyslexia and other learning needs. While the law says this data will help provide better services, it actually takes up so much time that teachers have less time to help students directly. The effort spent on recording information ends up taking away from teaching.  Consider for one moment the enormous amounts of data that accumulate in state and federal government databases.


#### More Loss of Local Control (are you detecting a pattern yet?)

The law claims that regional programs will bring services closer to home. But these "local" programs must still follow state rules, use state money, and meet state standards. While services may be delivered locally, the decisions are made by state officials who don’t know the students, leaving local educators with little control over the services they provide.


#### The Expert Time Sink

The law puts more money into special education evaluations and expert services to make the field more professional. However, the rules and paperwork requirements are so heavy that experts spend more time on administrative tasks than using their skills to help students directly.


#### The Resource Drain

The law promises large increases in special education funding, but the cost of managing all the new rules and reports is very likely to eat up more money than schools actually get for helping students. What seems like extra funding might not go as far as it seems.


### The Core Problem

These issues show that the law doesn’t understand how school systems improve. Real progress in schools happens through good relationships, local knowledge, and the flexibility to alter policies to better meet students' needs. Instead, this law focuses on complicated funding systems and state control that will get in the way of actual improvement.

<br>
<br>

## Detailed Analysis

### Main Table

```{r}
#| label: main-analysis-table-article4-sections50-65-revised

# Create analysis with new 7-code framework for Article 4 Sections 4.50-4.65
article4_likely_results_revised <- data.frame(
  Category = c(
    "Regional Day School Legislative Intent",
    "Regional Day School Statewide Plan",
    "School Offense Complaint Requirements",
    "School Offense Complaint Requirements",
    "Dyslexia Program Auditing Requirements",
    "Dyslexia Program Auditing Requirements", 
    "Dyslexia Program Auditing Requirements",
    "Dyslexia Program Auditing Requirements",
    "Dyslexia Program Auditing Requirements",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "PEIMS Reporting Requirements for Special Populations",
    "Special Education Funding Formula Changes",
    "Special Education Funding Formula Changes",
    "Special Education Funding Formula Changes",
    "Special Education Funding Formula Changes",
    "Special Education Service Group Allotment",
    "Special Education Service Group Allotment",
    "Special Education Service Group Allotment",
    "Special Education Service Group Allotment",
    "Special Education Transition Funding",
    "Special Education Transition Funding",
    "Special Education Transition Funding",
    "Dyslexia Funding Eligibility Changes",
    "Dyslexia Funding Eligibility Changes",
    "Dyslexia Funding Eligibility Changes",
    "College Readiness Outcomes Bonus Adjustments",
    "Special Transportation Rate Increases",
    "Special Education Evaluation Funding",
    "Special Education Evaluation Funding",
    "Foundation School Program Grant Allocation",
    "Federal Funds Distribution Requirements",
    "Day Placement Program Funding",
    "Day Placement Program Funding",
    "Day Placement Program Funding",
    "Day Placement Program Funding",
    "Parent-Directed Services Grants",
    "Parent-Directed Services Grants",
    "Parent-Directed Services Grants",
    "Regional Day School Program Funding"
  ),

  Incentive_Created = c(
    "Legislature intends to continue process of providing statewide suitable education to deaf/hard of hearing students under 22 years",
    "Director of services must develop and administer comprehensive statewide plan for educational services for deaf/hard of hearing students",
    "School offense complaints must be sworn by person with personal knowledge of underlying facts giving probable cause",
    "Complaints must be accompanied by statement from school employee stating whether child is eligible for or receives special education services",
    "Agency by rule must develop procedures to effectively audit and monitor and periodically conduct site visits of all school districts",
    "Agency must identify problems school districts experience in complying with dyslexia program requirements",
    "Agency must develop reasonable and appropriate remedial strategies to address school district noncompliance",
    "Agency must solicit input from parents of students enrolled in district during auditing and monitoring",
    "Agency must engage in general supervision activities including comprehensive system for monitoring described by Section 29.010",
    "Commissioner by rule must require districts to report number of students enrolled who are identified as having dyslexia",
    "Districts must report availability of school counselors including number of full-time equivalent counselors at each campus",
    "Districts must report availability of expanded learning opportunities as described by Section 33.252 at each campus",
    "Districts must report total number of students with whom district used intervention strategies at any time during year",
    "Districts must report total number of students to whom district provided aids accommodations or services under Section 504",
    "Districts must report disaggregated by campus and grade number of children who fail to attend school without excuse",
    "Districts must report number of students for whom district initiates truancy prevention measure under Section 25.0915",
    "Districts must report students enrolled in special education program under Subchapter A Chapter 29 as necessary for agency supervision",
    "For each student in average daily attendance in special education program district entitled to annual allotment multiplied by weight set by legislature",
    "Commissioner by rule must define eight tiers of intensity of service for use in determining funding under this section",
    "Commissioner must include one tier specifically addressing students receiving special education services in residential placement",
    "At least 55 percent of funds allocated under this section must be used in special education program under Subchapter A Chapter 29",
    "For each student in special education program district entitled to allotment in amount set by legislature for service group",
    "Commissioner by rule must establish at least four service groups for use in determining funding under this section",
    "Commissioner must consider type frequency and nature of services provided to student in establishing service groups",
    "At least 55 percent of funds allocated under service group allotment must be used for special education program",
    "For 2026-2027 school year commissioner may adjust weights or amounts to ensure compliance with federal maintenance requirements",
    "Commissioner must determine formulas through which school districts receive funding under Sections 48.102 and 48.1021",
    "Agency must provide technical assistance to school districts to ensure successful transition in funding formulas",
    "School district entitled to allotment only for student receiving instruction services or accommodations for dyslexia in accordance with individualized education program",
    "School district entitled to allotment for student receiving accommodations for dyslexia in accordance with plan developed under Section 504",
    "School district may use allotment to contract with private provider to provide supplemental academic services recommended under student program",
    "For annual graduate who demonstrates college career or military readiness and is enrolled in special education program district entitled to outcomes bonus of $4000",
    "School district or county providing special transportation services for eligible special education students entitled to state allocation at enhanced rate per mile",
    "For each child for whom school district conducts full individual and initial evaluation district entitled to allotment of $1000 or greater amount",
    "For 2025-2026 and 2026-2027 school years amount of allotment is $3000 for each child not enrolled in public school",
    "If commissioner determines amount appropriated for Foundation School Program exceeds amount to which districts entitled commissioner may provide grants",
    "After commissioner replaces withheld federal funds commissioner must distribute remaining amount to proportionately increase special education funding",
    "For each qualifying day placement program or cooperative that regional education service center or district establishes entitled to allotment up to $250000 first year",
    "Day placement program or cooperative entitled to $100000 each year after first year plus $150000 if at least three students enrolled",
    "Agency may not designate more than one day placement program or cooperative for service per county each year",
    "Agency may not provide allotment to more than 20 day placement programs or cooperatives for a year",
    "Student to whom agency awards grant under Subchapter A-1 Chapter 29 entitled to receive amount of $1500 or greater amount provided by appropriation",
    "Regional education service center designated to administer program entitled to amount equal to four percent of each grant awarded",
    "Total amount provided under parent-directed services grants may not exceed $80 million per school year with temporary increase to $150 million for 2025-2026",
    "Program administrator or fiscal agent of regional day school program for deaf entitled to receive allotment of $6925 per student receiving services"
  ),

  Intended_Consequence = c(
    "Continue legislative commitment to providing appropriate education for deaf and hard of hearing students",
    "Ensure systematic statewide approach to educational service delivery for students with hearing impairments",
    "Ensure proper documentation and due process protections for students involved in school discipline proceedings",
    "Provide essential information about special education status to inform appropriate handling of disciplinary matters",
    "Ensure compliance with state-mandated dyslexia program requirements through systematic monitoring and oversight",
    "Identify and address implementation challenges in dyslexia programming to improve service delivery",
    "Develop targeted interventions to address noncompliance and ensure program effectiveness",
    "Incorporate parent feedback into monitoring process to improve program quality and stakeholder satisfaction",
    "Ensure comprehensive oversight of special education compliance through integrated monitoring systems",
    "Create data collection system for tracking dyslexia identification and service provision across state",
    "Monitor availability of counseling support services at campus level to ensure adequate student support",
    "Track availability of extended learning opportunities to ensure equitable access across campuses",
    "Monitor intervention strategy usage to assess response to intervention implementation and effectiveness",
    "Track Section 504 service provision to ensure compliance with federal disability accommodation requirements",
    "Monitor chronic absenteeism patterns to identify students at risk and implement appropriate interventions",
    "Track truancy prevention efforts to ensure proactive approach to attendance issues",
    "Provide data necessary for agency to perform federal supervision requirements and determine appropriate funding levels",
    "Provide flexible funding system that accounts for varying levels of service intensity required by students with disabilities",
    "Create systematic approach to categorizing service intensity levels for more accurate funding allocation",
    "Ensure adequate funding for students requiring most intensive residential placement services",
    "Ensure majority of special education funding is used directly for special education programming rather than administrative purposes",
    "Implement alternative funding approach based on service groups rather than traditional placement-based model",
    "Create systematic categorization of services to better align funding with actual service delivery models",
    "Ensure funding allocation reflects actual service requirements and resource needs of individual students",
    "Maintain focus on direct special education service provision rather than administrative or indirect costs",
    "Provide transition mechanism to ensure federal compliance during implementation of new funding formulas",
    "Create unified funding approach that combines different special education funding streams for administrative efficiency",
    "Support districts during transition to new funding models to prevent service disruptions",
    "Focus dyslexia funding on students receiving formal educational interventions through IEP or 504 plan processes",
    "Expand eligibility for dyslexia funding to include students served through Section 504 accommodation plans",
    "Allow districts flexibility to contract for specialized dyslexia services while maintaining student school attendance",
    "Provide enhanced financial incentive for districts to support special education students in achieving post-secondary readiness",
    "Increase funding support for specialized transportation services required by students with disabilities",
    "Provide funding support for comprehensive evaluation processes required under federal special education law",
    "Provide enhanced funding for evaluation of children not currently enrolled in public school systems",
    "Allow flexible use of excess Foundation School Program funds to support special education monitoring and equipment needs",
    "Ensure special education programs receive proportionate share of available federal funding after compliance obligations met",
    "Provide startup and ongoing funding support for regional day placement programs serving students with intensive needs",
    "Create financial incentives for establishing day placement programs while providing ongoing operational support",
    "Prevent oversaturation of day placement programs within individual counties to ensure strategic resource allocation",
    "Control total number of day placement programs to ensure quality oversight and sustainable funding",
    "Provide direct funding to families for parent-directed special education services as alternative to traditional district-provided services",
    "Create administrative support structure for parent-directed services program through regional education service centers",
    "Establish substantial funding pool for parent-directed services while allowing for temporary expansion during implementation",
    "Ensure adequate and predictable funding for regional day school programs serving deaf and hard of hearing students"
  ),

  Likely_Result = c(
    "C, E, F", # Legislative intent: state dependency, admin burden, reduced local control
    "C, E, F", # Statewide plan: state dependency, admin burden, reduced local control
    "C, E, F", # Complaint requirements: state dependency, admin burden, reduced local control
    "C, E, F", # Employee statements: state dependency, admin burden, reduced local control
    "C, D, E, F", # Auditing procedures: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Problem identification: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Remedial strategies: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Parent input solicitation: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # General supervision: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Dyslexia reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Counselor reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Learning opportunities reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Intervention reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Section 504 reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Attendance reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Truancy reporting: state dependency, accountability substitute, admin burden, reduced local control
    "C, D, E, F", # Special education reporting: state dependency, accountability substitute, admin burden, reduced local control
    "A, B, C, E, F", # Funding formula: extrinsic motivation, competition, state dependency, admin burden, reduced local control
    "C, E, F", # Tier definitions: state dependency, admin burden, reduced local control
    "C, E, F", # Residential placement tier: state dependency, admin burden, reduced local control
    "C, E, F", # Fund usage requirements: state dependency, admin burden, reduced local control
    "A, B, C, E, F", # Service group allotment: extrinsic motivation, competition, state dependency, admin burden, reduced local control
    "C, E, F", # Service group establishment: state dependency, admin burden, reduced local control
    "C, E, F", # Service considerations: state dependency, admin burden, reduced local control
    "C, E, F", # Service group fund usage: state dependency, admin burden, reduced local control
    "C, E, F", # Transition funding adjustments: state dependency, admin burden, reduced local control
    "C, E, F", # Formula determination: state dependency, admin burden, reduced local control
    "C, E, F", # Technical assistance: state dependency, admin burden, reduced local control
    "C, E, F", # Dyslexia IEP eligibility: state dependency, admin burden, reduced local control
    "C, E, F", # Section 504 eligibility: state dependency, admin burden, reduced local control
    "A, C, E, F", # Private provider contracting: extrinsic motivation, state dependency, admin burden, reduced local control
    "A, B, C, E, F", # Outcomes bonus increase: extrinsic motivation, competition, state dependency, admin burden, reduced local control
    "A, C, E, F", # Transportation rate increase: extrinsic motivation, state dependency, admin burden, reduced local control
    "A, C, E, F", # Evaluation funding: extrinsic motivation, state dependency, admin burden, reduced local control
    "A, C, E, F", # Enhanced evaluation funding: extrinsic motivation, state dependency, admin burden, reduced local control
    "C, E, F", # Foundation program grants: state dependency, admin burden, reduced local control
    "C, E, F", # Federal funds distribution: state dependency, admin burden, reduced local control
    "A, B, C, E, F", # Day placement start up funding: extrinsic motivation, competition, state dependency, admin burden, reduced local control
    "A, B, C, E, F", # Day placement ongoing funding: extrinsic motivation, competition, state dependency, admin burden, reduced local control
    "C, E, F", # County designation limits: state dependency, admin burden, reduced local control
    "C, E, F", # Program number limits: state dependency, admin burden, reduced local control
    "A, C, E, F", # Parent-directed grants: extrinsic motivation, state dependency, admin burden, reduced local control
    "C, E, F", # Administrative percentage: state dependency, admin burden, reduced local control
    "C, E, F", # Grant funding limits: state dependency, admin burden, reduced local control
    "A, C, E, F" # Regional day school funding: extrinsic motivation, state dependency, admin burden, reduced local control
  )
)

# Verify all vectors have same length
stopifnot(length(article4_likely_results_revised$Category) == length(article4_likely_results_revised$Incentive_Created))
stopifnot(length(article4_likely_results_revised$Incentive_Created) == length(article4_likely_results_revised$Intended_Consequence))
stopifnot(length(article4_likely_results_revised$Intended_Consequence) == length(article4_likely_results_revised$Likely_Result))

# Create the main analysis table with revised codes
article4_main_table_revised <- datatable(
  article4_likely_results_revised,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '150px', targets = 0),   # Category  
      list(width = '300px', targets = 1),   # Incentive Created
      list(width = '250px', targets = 2),   # Intended Result
      list(width = '150px', targets = 3)    # Coded Results
    ),
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover',
  colnames = c('Category', 'Incentive Created', 'Intended Result', 'Codes for Destructive Results We Will Actually Get')
) %>%
  formatStyle(
    'Category',
    backgroundColor = styleEqual(
      c("Regional Day School Legislative Intent", "Regional Day School Statewide Plan", "School Offense Complaint Requirements", 
        "Dyslexia Program Auditing Requirements", "PEIMS Reporting Requirements for Special Populations", "Special Education Funding Formula Changes",
        "Special Education Service Group Allotment", "Special Education Transition Funding", "Dyslexia Funding Eligibility Changes",
        "College Readiness Outcomes Bonus Adjustments", "Special Transportation Rate Increases", "Special Education Evaluation Funding",
        "Foundation School Program Grant Allocation", "Federal Funds Distribution Requirements", "Day Placement Program Funding",
        "Parent-Directed Services Grants", "Regional Day School Program Funding"),
      c('#FFE8E8', '#FFD0D0', '#E8F4FD', '#E8F6FF', '#FFF2CC', '#E8F5E8', '#F0E8FF', '#FFE8F5', 
        '#FFEBCD', '#F0FFF0', '#F5F5DC', '#FFFACD', '#F0F8FF', '#FDF5E6', '#FAF0E6', '#FFF8DC', '#F5FFFA')
    ),
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Likely_Result',
    fontWeight = 'bold',
    fontSize = '14px',
    textAlign = 'center',
    backgroundColor = '#FFF8DC'
  )

# Display article 4 main table
article4_main_table_revised
```

<br>
<br>
<br>

### Supplemental Tables

```{r}
#| label: supplemental-analysis-article4-sections50-65-revised

# Create category analysis
category_analysis_a4_revised <- article4_likely_results_revised %>%
  group_by(Category) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(desc(Count)) %>%
  rename(`Incentive Category` = Category,
         `Number of Incentives` = Count)

# Analyze most frequent negative result codes
code_frequency_a4_revised <- article4_likely_results_revised %>%
  separate_rows(Likely_Result, sep = ", ") %>%
  count(Likely_Result, sort = TRUE) %>%
  rename(`Negative Result Code` = Likely_Result,
         `Frequency` = n) %>%
  mutate(Percentage = round(100 * Frequency / nrow(article4_likely_results_revised), 3))

# Key policy findings focused on coded patterns
key_findings_a4_revised <- data.frame(
  Finding = c(
    "Most Universal Negative Patterns",
    "State Dependency Complete Takeover", 
    "Administrative Burden Universal Impact",
    "Local Control Elimination Pattern",
    "Accountability Substitution Pattern",
    "Most Problematic Policy Areas",
    "Extrinsic Motivation Corruption Pattern", 
    "Competition Over Cooperation Pattern"
  ),
  Description = c(
    "Codes C (state dependency), E (admin burden), and F (local control loss) appear in 100.000% of incentives, indicating universal centralization and bureaucratization",
    "Code C (state dependency) appears in 100.000% of incentives, showing complete shift from local autonomy to state control in specialized education",
    "Code E (admin burden) appears in 100.000% of incentives, creating massive compliance overhead across all special education funding and reporting policies",
    "Code F (local control loss) appears in 100.000% of incentives, representing complete elimination of district flexibility in specialized education decisions",
    "Code D (accountability substitute) appears in 37.778% of incentives, concentrated in monitoring and reporting requirements that replace system improvement",
    "PEIMS Reporting Requirements and Special Education Funding Formula Changes show highest concentrations with 8 and 4 harmful incentives respectively",
    "Code A (extrinsic motivation) appears in 31.111% of incentives, systematically undermining professional dedication through financial incentives", 
    "Code B (competition over cooperation) appears in 13.333% of incentives, creating resource competition rather than collaborative improvement"
  )
)

# === DISPLAY SUPPLEMENTAL TABLES ===

# Category Analysis Header
datatable(
  data.frame(Title = "INCENTIVE CATEGORY ANALYSIS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#FF9800',
              color = 'white')

# Category Analysis Table
datatable(
  category_analysis_a4_revised,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Number of Incentives',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(category_analysis_a4_revised$`Number of Incentives`), '#E8F4FD')) %>%
  formatStyle('Incentive Category',
              fontWeight = 'bold',
              textAlign = 'left')

# Code Frequency Analysis Header
datatable(
  data.frame(Title = "NEGATIVE RESULT CODE FREQUENCY"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#DC143C',
              color = 'white')

# Code Frequency Table
datatable(
  code_frequency_a4_revised,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Frequency',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(code_frequency_a4_revised$Frequency), '#FFB6C1')) %>%
  formatStyle('Negative Result Code',
              fontWeight = 'bold',
              textAlign = 'center',
              fontSize = '16px') %>%
  formatStyle('Percentage',
              fontWeight = 'bold',
              textAlign = 'center')

# Key Findings Header
datatable(
  data.frame(Title = "KEY CODED PATTERN FINDINGS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#F44336',
              color = 'white')

# Key Findings Table
datatable(
  key_findings_a4_revised,
  filter = 'top',
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '25%', targets = 0),
      list(width = '75%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Finding',
              fontWeight = 'bold',
              backgroundColor = '#FFE4E1',
              textAlign = 'left') %>%
  formatStyle('Description',
              textAlign = 'left')

# Create coded summary statistics
summary_statistics_a4_revised <- data.frame(
  Metric = c(
    "Total Incentives Analyzed",
    "Most Problematic Code", 
    "Most Frequent Code Combination",
    "Policy Categories with Code C",
    "Policy Categories with Code F",
    "Average Codes per Incentive"
  ),
  Value = c(
    nrow(article4_likely_results_revised),
    paste0("Codes C, E, F (all appear ", max(code_frequency_a4_revised$Frequency), " times each)"),
    "C, E, F (state dependency + admin burden + local control loss)",
    sum(grepl("C", article4_likely_results_revised$Likely_Result)),
    sum(grepl("F", article4_likely_results_revised$Likely_Result)),
    round(mean(lengths(strsplit(article4_likely_results_revised$Likely_Result, ", "))), 3)
  )
)

# Summary Statistics Header
datatable(
  data.frame(Title = "CODED ANALYSIS SUMMARY STATISTICS"),
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#9C27B0',
              color = 'white')

# Summary Statistics Table
datatable(
  summary_statistics_a4_revised,
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '40%', targets = 0),
      list(width = '60%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Metric',
              fontWeight = 'bold',
              textAlign = 'left') %>%
  formatStyle('Value',
              textAlign = 'left',
              fontWeight = 'bold')
```

<br>
<br>

### Key List of Likely Negative Results 

#### The evaluation codes are based on these axioms:

- Central planning (one size fits all) degrades everyone's performance
  - Makes nearly every district action a legal matter
  - Robs people of pride in their work
  
- Efforts at 'accountability' will not improve performance of anyone

- Cooperation within an organization is far more effective than competition 

- All incentives work; some promote and some pervert the ***intention*** of the incentive


```{r}
#| label: negative-results-coding-key-article4-sections50-65-revised

# Create negative results table with color coding for new 7-code framework
negative_results_a4_revised <- data.frame(
  Code = c("A", "B", "C", "D", "E", "F", "G"),

  Likely_Negative_Result = c(
    "Decreases intrinsic motivation in favor of extrinsic motivation",
    "Decreases cooperation within districts in favor of competition for resources",  
    "Increases dependency on State funding and bureaucracy",
    "Substitutes 'accountability' for improvement of the system people work in",
    "Increases district administrative burdens",
    "Decreases local control and flexibility",
    "Treats people as the 'problem', instead of the system they work in"
  )
)

# Define bright contrasting color palette (7 colors)
color_palette_a4_revised <- c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF")

# Create color assignments for each row
row_colors_a4_revised <- color_palette_a4_revised[1:nrow(negative_results_a4_revised)]

# Display the coding key table
DT::datatable(
  negative_results_a4_revised,
  options = list(
    pageLength = 15,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    columnDefs = list(
      list(width = '40px', targets = 0),  # Narrow first column
      list(width = '90%', targets = 1)    # Wide second column
    ),
    scrollX = FALSE,
    autoWidth = FALSE
  ),
  rownames = FALSE,
  class = 'display'
) %>%
  formatStyle(
    'Code',
    textAlign = 'right',
    fontWeight = 'bold',
    backgroundColor = styleEqual(negative_results_a4_revised$Code, row_colors_a4_revised),
    color = 'white'
  ) %>%
  formatStyle(
    'Likely_Negative_Result',
    backgroundColor = styleEqual(negative_results_a4_revised$Code, row_colors_a4_revised),
    color = 'black'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    border = '2px solid #333',
    borderCollapse = 'collapse'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    maxWidth = '9in',
    tableLayout = 'fixed'
  )

```


