---
title: "Texas HB2 Article 4: Special Education (Sections 4.01-4.15) - Incentives and Likely Results"
author: "Dan Swart and Claude Sonnet 4.0"
description: "Analysis covers Article 4 (Special Education) only - Sections 4.01 through 4.15. Analysis focuses on special education agency duties, compliance monitoring, eligibility criteria, interagency coordination, and residential placement oversight. Document spans pages 25-31 of legislative text."
format:
    html:
      toc: true
      toc-float: true
      page-layout: article
      embed-resources: true
      code-fold: true
      code-summary: "Show the code"
      code-tools: true
      code-overflow: wrap
      code-block-bg: "#FAEBD7"
      code-block-border-left: "#31BAE9"
      code-link: true
      fig-width: 8
      fig-height: 6
      fig-align: center
      html-math-method: katex
      css: swart-20250327.css
    pdf:
      toc: true
      number-sections: true
      colorlinks: true
      papersize: letter
      geometry:
        - margin=1in
      fig-width: 12
      fig-height: 10
      fig-pos: 'H'
    typst:
      toc: true
      fig-width: 12
      fig-height: 10
      keep-tex: true
      prefer-html: true
---

```{r}
#| label: setup-article4-sections01-15
#| include: false
#| echo: false
#| message: false
#| warning: false

library(tidyverse)  # For dplyr, ggplot, and friends
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(DiagrammeR) # For complete control of the layout
library(knitr)      # For controlling rendering
library(kableExtra) # For tables summarizing results
library(DT)         # For rendering content that kableExtra cannot
library(dplyr)      # For data wrangling
```

<br>
<br>

## Opinionated Executive Summary and Counter-Intuitive Implications

The legislative code analyzed in Article 4's special education framework reveals a systematic dismantling of local educational judgment disguised as "student protection" and "federal compliance." All code sections are evaluated based on my own "Table of Codes for Destructive Results We Will Actually Get" (see Table below).

Most disturbing is that 100% of incentives contain Code C (failure by force) and Code G (reduced local control), indicating complete state takeover of decisions that should belong to educators who actually know the children.

#### The Student Support-Professional Disempowerment Paradox:

Visual impairment and hearing services expansion promises comprehensive support for students with sensory disabilities through systematic state agency coordination, yet my coding reveals these policies will create forced failure (Code C in 100% of incentives) while eliminating local flexibility to respond to individual student needs. Like requiring all restaurants to serve identical meals regardless of customer preferences, these policies assume bureaucratic standardization improves outcomes better than professional relationships between teachers, families, and students.

#### The Federal Compliance-Local Chaos Multiplication:

IEP committee composition and eligibility criteria updates appear to align state practices with federal requirements through clear procedural guidelines, yet the analysis shows systematic tampering (Code I in 100% of incentives) that will destabilize existing functional special education teams. Districts that have spent years building trust and expertise around individual students will be forced to reconstruct their entire approach to serve compliance monitoring rather than children, creating the educational equivalent of requiring surgeons to get permission from accountants before operating.

#### The Comprehensive System-Service Fragmentation Inversion:

The Special Education Implementation Law Restructure promises seamless statewide service delivery through coordinated regional support systems, yet shows the highest concentration of negative codes with 9 separate harmful incentives including complete bureaucratic burden creation (Code L in 100% of cases). Instead of improving services, this creates an elaborate administrative machinery that consumes resources while delaying actual help to students who need immediate support.

#### The Quality Assurance-Resource Waste Amplification:

Monitoring system overhauls and progressive sanctions promise accountability through comprehensive oversight and targeted interventions, yet Code M (resource waste) appears in 22.5% of incentives, particularly concentrated in the very programs claiming to improve efficiency. Districts will spend millions documenting compliance for evaluators who have never worked directly with students with disabilities, creating the special education equivalent of requiring teachers to file environmental impact reports before helping struggling readers.

#### The Parent Empowerment-Communication Theater Contradiction:

Language translation requirements and information material systems appear to remove barriers for non-English speaking families through systematic translation services and comprehensive placement information, yet these create bureaucratic burdens (Code L) that will actually delay communication rather than improve it. Like requiring hospitals to translate every medical conversation into 47 languages before treating patients, these policies prioritize process documentation over meaningful family engagement.

#### The Personnel Shortage-Training Bureaucracy Multiplication:

Regional service center support mandates and technical assistance requirements promise to address the special education personnel crisis through coordinated training and recruitment strategies, yet they systematically create system tampering (Code I in 100% of cases) that will drive experienced professionals out of the field. Instead of supporting teachers who work with students with disabilities, these policies create compliance specialist positions that pay more than direct service roles, accelerating the very personnel crisis they claim to address.

#### The Individual Focus-System Standardization Paradox:

Special education diploma achievement requirements and ARD committee decision-making authority appear to ensure individualized programming through flexible assessment options and team-based planning, yet Code F (individual vs system focus problems) appears in 17.5% of incentives, indicating these policies will force students into standardized pathways rather than truly individualized approaches. The ultimate irony: policies claiming to individualize education actually eliminate the local flexibility necessary for true individualization.

#### The Interagency Coordination-Administrative Multiplication Escalation:

Memorandum requirements and state living center information systems promise seamless coordination between education and human services agencies through clear role definitions and systematic information sharing, yet they create layers of bureaucratic interface that will slow rather than speed services to families navigating multiple systems simultaneously.

#### Summary

These contradictions reveal a special education system that has abandoned the development of professional judgment, family relationships, and individualized programming in favor of centralized compliance monitoring, procedural documentation, and bureaucratic coordination mechanisms.

Like trying to improve hospitals by requiring doctors to get approval from insurance companies before every medical decision, these policies assume students with disabilities are better served by regulatory oversight than by knowledgeable professionals working directly with families.

They systematically undermine the local decision-making authority, stable service relationships, and individualized programming flexibility that actually create effective special education outcomes.

Most devastatingly, they create a special education industrial complex that costs tens of millions in state funding while producing administrators and consultants trained to navigate bureaucratic systems rather than to build relationships with students, families, and communities around individual needs and strengths.

This ultimately harms the very children with disabilities these policies claim to protect, while enriching the compliance consultants, monitoring specialists, and oversight bureaucrats who benefit from the chaos—many of whom have never worked directly with a student with disabilities in their entire careers.
<br>
<br>

## Detailed Analysis

### Main Table

```{r}
#| label: main-analysis-table-article4-sections01-15

# Create analysis with coded negative results for Article 4 Sections 4.01-4.15
article4_likely_results <- data.frame(
  Category = c(
    "Visual Impairment Agency Duties Expansion",
    "Visual Impairment Agency Duties Expansion",
    "Special Supplemental Allowance Distribution System",
    "Regional Service Center Core Services Mandates",
    "Regional Service Center Core Services Mandates",
    "Regional Service Center Core Services Mandates",
    "Special Education Diploma Achievement Requirements",
    "Special Education Diploma Achievement Requirements",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Special Education Implementation Law Restructure",
    "Annual Special Education Board Meeting Requirements",
    "Annual Special Education Board Meeting Requirements",
    "Special Education Eligibility Criteria Updates",
    "Special Education Eligibility Criteria Updates",
    "Special Education Eligibility Criteria Updates",
    "IEP Committee Composition Federal Alignment",
    "IEP Committee Composition Federal Alignment",
    "Parent Language Translation Requirements",
    "Parent Language Translation Requirements",
    "Commissioner Autism Information Authority",
    "State Living Center Information Materials System",
    "State Living Center Information Materials System",
    "State Living Center Information Materials System",
    "Special Education Advisory Committee Restructure",
    "Special Education Advisory Committee Restructure",
    "Special Education Contract Approval System",
    "Special Education Contract Approval System",
    "Special Education Monitoring System Overhaul",
    "Special Education Monitoring System Overhaul",
    "Special Education Progressive Sanctions System",
    "Special Education Progressive Sanctions System",
    "Special Education Complaint Management System",
    "Interagency Coordination Memorandum Requirements"
  ),

  Incentive_Created = c(
    "Agency must carry out duties for children with visual impairments, who are deaf or hard of hearing, or who are deaf-blind",
    "Commissioner must develop system to distribute special supplemental allowances for students with visual impairments",
    "Regional education service centers must distribute special supplemental allowances as required under revised section reference",
    "Regional service centers must maintain core services including assistance for districts out of special education compliance",
    "Core services must include training for programs qualifying for funding allotments under specified sections",
    "Assistance must be specifically designed for districts assigned unacceptable performance ratings or compliance failures",
    "Special education students may earn distinguished achievement or endorsements by completing curriculum with or without modifications",
    "ARD committee must determine whether student must achieve satisfactory performance on end-of-course assessments",
    "Agency must develop comprehensive system to ensure statewide and local compliance with federal and state special education law",
    "Comprehensive system must include provision of services primarily through school districts and shared services arrangements",
    "System must focus on maximizing student outcomes and include rulemaking, technical assistance, and monitoring protocols",
    "System must facilitate interagency coordination when other state agencies deliver services to students with disabilities",
    "System must pursue strategies to meet statewide special education and related services personnel needs",
    "System must ensure regional education service centers maintain regional support functions throughout the state",
    "System must effectively monitor and conduct site visits to ensure rules are applied consistently",
    "System must provide training and technical assistance to ensure appropriately trained personnel serve students",
    "System must ensure individualized education programs are properly developed and implemented in least restrictive environment",
    "School district and charter school boards must include annual discussion of special education student performance",
    "Agency must adopt performance indicators for measuring quality of learning and achievement for special education students",
    "Agency must develop specific eligibility criteria based on general classifications and federal law requirements",
    "Students eligible from birth through 21 years if visual impairment, deaf, hard of hearing, or deaf-blind prevents safe education",
    "Students from 3-21 years eligible if disabilities prevent adequate or safe education without special education services",
    "IEP committees must include persons required under federal law with regular education teacher when appropriate",
    "Regular education teacher must be responsible for implementing portion of child's individualized education program",
    "Districts must provide Spanish translation of IEP if Spanish is parent's primary language",
    "Districts must make good faith effort to translate IEP into parent's primary language if not Spanish",
    "Commissioner may require inclusion of autism or pervasive developmental disorder information in student's IEP",
    "Health and Human Services Commission must develop materials on educational residential placement options",
    "Materials must include information on full continuum of educational residential placement options for qualifying children",
    "Agency must make materials available to school districts for distribution to parents at ARD committee meetings",
    "Governor must appoint continuing advisory committee consistent with federal requirements",
    "Advisory committee members appointed for staggered four-year terms with specified member qualifications",
    "Commissioner must set minimum standards and develop list of approved facilities for student services",
    "Commissioner may approve facilities only after programmatic evaluation of qualifications, costs, and curriculum content",
    "Agency must develop comprehensive monitoring system including cyclical process and targeted risk-based process",
    "Agency may require districts to obtain specialized technical assistance for documented noncompliance issues",
    "Agency must develop system of interventions and sanctions for districts identified as being in noncompliance",
    "Sanctions must range in severity and may include withholding funds with alternative service arrangements",
    "Agency must develop expedited investigation and resolution system for special education service failure complaints",
    "Agencies must develop memorandum establishing respective responsibilities for students with disabilities in residential facilities"
  ),

  Intended_Consequence = c(
    "Expand state support to include broader population of students with hearing impairments in addition to visual impairments",
    "Ensure adequate financial support reaches districts serving students with various sensory impairments through systematic distribution",
    "Maintain consistent funding distribution while updating administrative references to reflect current legal framework",
    "Provide targeted technical assistance to districts struggling with special education compliance requirements",
    "Guarantee districts receive appropriate support for implementing special education programs that qualify for state funding",
    "Focus intensive assistance on districts with documented performance problems or compliance failures",
    "Enable special education students to achieve high academic recognition while accommodating individual learning needs",
    "Ensure individualized decision-making about assessment requirements based on each student's capabilities and needs",
    "Create systematic statewide approach to ensure consistent special education compliance across all districts",
    "Establish clear service delivery framework using existing district structures supplemented by regional support",
    "Shift focus from procedural compliance to actual student outcomes while maintaining necessary oversight mechanisms",
    "Improve coordination between education agencies and health/human services agencies serving students with disabilities",
    "Address statewide shortage of qualified special education personnel through strategic recruitment and training initiatives",
    "Ensure regional service centers can effectively assist school districts in identifying and developing appropriate services",
    "Maintain consistent application of special education rules across all districts through systematic monitoring and site visits",
    "Guarantee that qualified personnel with appropriate training work directly with students receiving special education services",
    "Ensure every student receives individualized programming appropriate to their needs in least restrictive educational setting",
    "Increase transparency and public accountability for special education program performance at local district level",
    "Establish measurable standards for evaluating quality and effectiveness of special education services and student achievement",
    "Update eligibility determination processes to reflect current federal requirements and contemporary diagnostic practices",
    "Ensure students with sensory impairments receive appropriate special education services when regular education is insufficient",
    "Provide clear criteria for determining when students with various disabilities require special education services",
    "Align IEP committee composition with federal requirements while ensuring appropriate educator participation",
    "Guarantee that regular education teachers involved in IEP development have actual implementation responsibilities",
    "Remove language barriers for Spanish-speaking parents to ensure meaningful participation in educational planning",
    "Extend language accessibility to parents speaking languages other than Spanish for equal participation opportunities",
    "Ensure students with autism receive comprehensive individualized programming necessary for appropriate public education",
    "Provide parents with complete information about residential placement options when considering educational alternatives",
    "Ensure families understand full range of residential educational options available through systematic information sharing",
    "Make comprehensive placement information readily available to parents through school districts during ARD meetings",
    "Maintain federal compliance for advisory committee structure while ensuring representation from special education professionals",
    "Provide continuity and expertise through staggered appointments and inclusion of experienced special education administrators",
    "Establish quality assurance for facilities serving students with disabilities through rigorous standards and evaluation",
    "Protect students by ensuring only qualified facilities meeting high standards receive approval for service provision",
    "Create comprehensive monitoring system addressing both compliance requirements and educational quality indicators",
    "Provide targeted support to districts experiencing compliance difficulties or showing data indicators of service problems",
    "Establish meaningful consequences for continued noncompliance while maintaining focus on student service delivery",
    "Create escalating intervention system that can redirect resources to ensure continued services despite district failures",
    "Provide efficient resolution mechanism for parents experiencing delays or inadequate special education services",
    "Clarify responsibilities and improve coordination between agencies serving students with disabilities in residential settings"
  ),

  Likely_Result = c(
    "C, G, I, L", # Visual impairment expansion: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Allowance distribution: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Service center distribution: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Core services compliance: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Training assistance: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Performance assistance: failure by force, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Diploma requirements: failure by force, individual focus, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Assessment decisions: failure by force, individual focus, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Comprehensive system: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Service delivery framework: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Student outcomes focus: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Interagency coordination: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Personnel strategies: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Regional support functions: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Monitoring and site visits: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Personnel training: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # IEP development: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Annual board meetings: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Performance indicators: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Eligibility criteria: failure by force, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Sensory impairment eligibility: failure by force, individual focus, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Disability criteria: failure by force, individual focus, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # IEP committee composition: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Teacher responsibilities: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Spanish translation: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Language translation: failure by force, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Autism information: failure by force, individual focus, reduced local control, system tampering, bureaucracy
    "C, G, I, K, L, M", # Living center materials: failure by force, reduced local control, system tampering, tax burden, bureaucracy, waste resources
    "C, G, I, L", # Placement information: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Material distribution: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Advisory committee: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Committee appointments: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Facility standards: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Facility approval: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Monitoring system: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Technical assistance: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Sanctions system: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Intervention escalation: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Complaint system: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L" # Memorandum requirements: failure by force, reduced local control, system tampering, bureaucracy
  )
)

# Create the main analysis table with coded results
article4_main_table <- datatable(
  article4_likely_results,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '150px', targets = 0),   # Category  
      list(width = '300px', targets = 1),   # Incentive Created
      list(width = '250px', targets = 2),   # Intended Result
      list(width = '150px', targets = 3)    # Coded Results
    ),
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover',
  colnames = c('Category', 'Incentive Created', 'Intended Result', 'Codes for Destructive Results We Will Actually Get')
) %>%
  formatStyle(
    'Category',
    backgroundColor = styleEqual(
      c("Visual Impairment Agency Duties Expansion", "Special Supplemental Allowance Distribution System", "Regional Service Center Core Services Mandates", 
        "Special Education Diploma Achievement Requirements", "Special Education Implementation Law Restructure", "Annual Special Education Board Meeting Requirements",
        "Special Education Eligibility Criteria Updates", "IEP Committee Composition Federal Alignment", "Parent Language Translation Requirements",
        "Commissioner Autism Information Authority", "State Living Center Information Materials System", "Special Education Advisory Committee Restructure",
        "Special Education Contract Approval System", "Special Education Monitoring System Overhaul", "Special Education Progressive Sanctions System",
        "Special Education Complaint Management System", "Interagency Coordination Memorandum Requirements"),
      c('#FFE8E8', '#FFD0D0', '#E8F4FD', '#E8F6FF', '#FFF2CC', '#E8F5E8', '#F0E8FF', '#FFE8F5', 
        '#FFE0B2', '#E0F0FF', '#F5F5DC', '#FFB6C1', '#FFE4E1', '#E8E8E8', '#F0F8FF', '#E6E6FA', '#FFF8DC')
    ),
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Likely_Result',
    fontWeight = 'bold',
    fontSize = '14px',
    textAlign = 'center',
    backgroundColor = '#FFF8DC'
  )

# Display article 4 main table
article4_main_table
```

<br>
<br>
<br>

### Supplemental Tables

```{r}
#| label: supplemental-analysis-article4-sections01-15

# Create category analysis
category_analysis_a4 <- article4_likely_results %>%
  group_by(Category) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(desc(Count)) %>%
  rename(`Incentive Category` = Category,
         `Number of Incentives` = Count)

# Analyze most frequent negative result codes
code_frequency_a4 <- article4_likely_results %>%
  separate_rows(Likely_Result, sep = ", ") %>%
  count(Likely_Result, sort = TRUE) %>%
  rename(`Negative Result Code` = Likely_Result,
         `Frequency` = n) %>%
  mutate(Percentage = round(100 * Frequency / nrow(article4_likely_results), 1))

# Key policy findings focused on coded patterns
key_findings_a4 <- data.frame(
  Finding = c(
    "Most Common Negative Pattern",
    "System Tampering and Control Dominance", 
    "Bureaucratic Implementation Pattern",
    "Local Control Reduction Pattern",
    "Individual vs System Focus Imbalance",
    "Most Problematic Policy Areas",
    "Resource Waste in Oversight Programs",
    "Financial Motivation Introduction"
  ),
  Description = c(
    "Code C (failure by force) appears in 100.0% of incentives, followed by Code G (reduced local control) at 100.0%",
    "System tampering affects every single incentive, indicating complete disruption of existing special education systems",
    "Administrative burdens appear in 100.0% of incentives, creating massive compliance overhead across all special education policies",
    "Code G (reduced local control) appears in 100.0% of incentives, representing complete state takeover of special education decisions", 
    "Code F (individual vs system focus) appears in 17.5% of incentives, primarily in student eligibility and assessment modifications",
    "Special Education Implementation Law shows highest concentration of negative codes with 9 separate harmful incentives created",
    "Code M (waste resources) appears in 22.5% of incentives, particularly in monitoring systems and board meeting requirements",
    "No financial motivation codes (A or B) appear, indicating focus on control rather than monetary incentives in special education"
  )
)

# === DISPLAY SUPPLEMENTAL TABLES ===

# Category Analysis Header
datatable(
  data.frame(Title = "INCENTIVE CATEGORY ANALYSIS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#FF9800',
              color = 'white')

# Category Analysis Table
datatable(
  category_analysis_a4,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Number of Incentives',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(category_analysis_a4$`Number of Incentives`), '#E8F4FD')) %>%
  formatStyle('Incentive Category',
              fontWeight = 'bold',
              textAlign = 'left')

# Code Frequency Analysis Header
datatable(
  data.frame(Title = "NEGATIVE RESULT CODE FREQUENCY"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#DC143C',
              color = 'white')

# Code Frequency Table
datatable(
  code_frequency_a4,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Frequency',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(code_frequency_a4$Frequency), '#FFB6C1')) %>%
  formatStyle('Negative Result Code',
              fontWeight = 'bold',
              textAlign = 'center',
              fontSize = '16px') %>%
  formatStyle('Percentage',
              fontWeight = 'bold',
              textAlign = 'center')

# Key Findings Header
datatable(
  data.frame(Title = "KEY CODED PATTERN FINDINGS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#F44336',
              color = 'white')

# Key Findings Table
datatable(
  key_findings_a4,
  filter = 'top',
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '25%', targets = 0),
      list(width = '75%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Finding',
              fontWeight = 'bold',
              backgroundColor = '#FFE4E1',
              textAlign = 'left') %>%
  formatStyle('Description',
              textAlign = 'left')

# Create coded summary statistics
summary_statistics_a4 <- data.frame(
  Metric = c(
    "Total Incentives Analyzed",
    "Most Problematic Code", 
    "Most Frequent Code Combination",
    "Policy Categories with Code C",
    "Policy Categories with Code G",
    "Average Codes per Incentive"
  ),
  Value = c(
    nrow(article4_likely_results),
    paste0("Code C and G (both appear ", max(code_frequency_a4$Frequency), " times each)"),
    "C, G, I, L (force + control + tampering + bureaucracy)",
    sum(grepl("C", article4_likely_results$Likely_Result)),
    sum(grepl("G", article4_likely_results$Likely_Result)),
    round(mean(lengths(strsplit(article4_likely_results$Likely_Result, ", "))), 1)
  )
)

# Summary Statistics Header
datatable(
  data.frame(Title = "CODED ANALYSIS SUMMARY STATISTICS"),
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#9C27B0',
              color = 'white')

# Summary Statistics Table
datatable(
  summary_statistics_a4,
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '40%', targets = 0),
      list(width = '60%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Metric',
              fontWeight = 'bold',
              textAlign = 'left') %>%
  formatStyle('Value',
              textAlign = 'left',
              fontWeight = 'bold')
```

<br>
<br>

### Key List of Likely Negative Results (Revised 7-Code Framework)

#### The evaluation codes are based on these axioms:

- Central planning (one size fits all) degrades everyone's performance
  - Makes nearly every district action a legal matter
  - Robs people of pride in their work
  
- Efforts at 'accountability' will not improve performance of anyone

- Cooperation within an organization is far more effective than competition 

- All incentives work; some promote and some pervert the ***intention*** of the incentive


```{r}
#| label: negative-results-coding-key-article4-sections01-15

# Create negative results table with color coding
negative_results_a4 <- data.frame(
  Code = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N"),

  Likely_Negative_Result = c(
    "Will move people from intrinsic to extrinsic motivation",
    "Will makes money the most important motivation",  
    "Will fail because improvements cannot be realized by force from the state",
    "People will become 'the problem', not the system that they work in",
    "Will decrease collaboration, pitting people against each other for rewards",
    "Will decrease opportunities and efforts to improve the system teachers work in; try to 'improve' individual teachers instead",
    "Will reduce local control",
    "Will reduce economic agility of districts when they need it most",
    "Will tamper with existing stable systems, making things worse",
    "Impedes career development when goal posts constantly moving",
    "Shifts the onus of tax increases to districts to pay for unfunded state mandates",
    "Will create more administrative and bureaucratic burdens",
    "Will cause administration and teachers to waste even more resources, while bankrupting financially responsible districts",
    "Will create long-term district obligations that are not sustainable when state funding is changed every 2 years"
  )
)

# Define bright contrasting color palette (7 colors)
color_palette_a4 <- c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF")

# Create color assignments for each row (cycling through 7 colors for 14 rows)
row_colors_a4 <- rep(color_palette_a4, length.out = nrow(negative_results_a4))

# Display the coding key table
DT::datatable(
  negative_results_a4,
  options = list(
    pageLength = 15,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    columnDefs = list(
      list(width = '40px', targets = 0),  # Narrow first column
      list(width = '90%', targets = 1)    # Wide second column
    ),
    scrollX = FALSE,
    autoWidth = FALSE
  ),
  rownames = FALSE,
  class = 'display'
) %>%
  formatStyle(
    'Code',
    textAlign = 'right',
    fontWeight = 'bold',
    backgroundColor = styleEqual(negative_results_a4$Code, row_colors_a4),
    color = 'white'
  ) %>%
  formatStyle(
    'Likely_Negative_Result',
    backgroundColor = styleEqual(negative_results_a4$Code, row_colors_a4),
    color = 'black'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    border = '2px solid #333',
    borderCollapse = 'collapse'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    maxWidth = '9in',
    tableLayout = 'fixed'
  )

```

