---
title: "Texas HB2 Article 5: Early Childhood Education (Sections 5.01-5.09) - Incentives and Likely Results"
author: "Dan Swart and Claude Sonnet 4.0"
description: "Analysis covers Article 5 (Early Childhood Education) Sections 5.01 through 5.09 only. Analysis focuses on open-enrollment charter school requirements, teacher literacy and mathematics achievement academies, school year planning grants, reading readiness assessments, early literacy instruments, and reading intervention programs. Document spans pages 55-59 of legislative text."
format:
    html:
      toc: true
      toc-float: true
      page-layout: article
      embed-resources: true
      code-fold: true
      code-summary: "Show the code"
      code-tools: true
      code-overflow: wrap
      code-block-bg: "#FAEBD7"
      code-block-border-left: "#31BAE9"
      code-link: true
      fig-width: 8
      fig-height: 6
      fig-align: center
      html-math-method: katex
      css: swart-20250327.css
    pdf:
      toc: true
      number-sections: true
      colorlinks: true
      papersize: letter
      geometry:
        - margin=1in
      fig-width: 12
      fig-height: 10
      fig-pos: 'H'
    typst:
      toc: true
      fig-width: 12
      fig-height: 10
      keep-tex: true
      prefer-html: true
---

```{r}
#| label: setup-article5-sections01-09
#| include: false
#| echo: false
#| message: false
#| warning: false

library(tidyverse)  # For dplyr, ggplot, and friends
library(ggdag)      # For plotting DAGs
library(dagitty)    # For working with DAG logic
library(DiagrammeR) # For complete control of the layout
library(knitr)      # For controlling rendering
library(kableExtra) # For tables summarizing results
library(DT)         # For rendering content that kableExtra cannot
library(dplyr)      # For data wrangling
```

<br>
<br>

## Opinionated Executive Summary:  Good Intentions, Unintended Results

The legislative code analyzed in Article 5, Sections 5.01-5.09 reveals several fascinating paradoxes that expose the fundamental contradictions between educational choice rhetoric and centralized control reality:

#### The Charter School Freedom Paradox:

Charter schools were originally created to escape the bureaucratic constraints of traditional public education, yet this legislation subjects them to identical state oversight requirements as district schools. The "choice" becomes purely cosmetic - families can choose between different buildings that operate under identical state mandates, eliminating the innovation and flexibility that justified charter schools' existence.

#### The Professional Development Professionalization Paradox:

Teacher literacy and mathematics academies appear to elevate and professionalize teaching through specialized training, yet the stipend system and mandatory evaluation requirements actually deprofessionalize the work by reducing complex pedagogical decisions to checklist completion and financial incentives rather than intrinsic educational purpose.

#### The Early Assessment Support-Surveillance Contradiction: 

Kindergarten reading readiness assessments are framed as supporting early identification of students needing help, yet the raw score reporting requirements to state agencies and integration into "school readiness certification systems" transform developmental assessment into surveillance data collection that may stigmatize children before they begin formal education.

#### The Grant Support-Control Inversion:

School year planning grants appear to provide helpful financial support for districts extending their academic calendar, yet the grants are specifically designed to help districts "qualify for incentive funding" - meaning the state pays districts to plan how to access more state money with more state strings attached, creating financial dependency rather than educational improvement.

#### The Reading Intervention Choice-Mandate Paradox: 

Students "assigned" to reading intervention programs must attend "unless specifically exempted," creating the illusion of individualized educational decisions while actually implementing one-size-fits-all mandates that override parent and teacher judgment about what individual children need.

#### The Cost-Free Assessment Expense Paradox:

Reading instruments are provided to districts "at no cost" to remove financial barriers, yet the administrative time, training requirements, data reporting obligations, and evaluation systems required to use these "free" instruments may cost districts more than purchasing their own assessments would have cost, while providing less useful information for actual instruction.

#### Summary

These contradictions reveal a system that has abandoned its stated educational goals in favor of administrative control mechanisms that create the appearance of improvement while systematically undermining the professional relationships, local decision-making, and educational flexibility that actually produce learning.

<br>

## Detailed Analysis

### Main Table

```{r}
#| label: main-analysis-table-article5-sections01-09
#| echo: false
#| warning: false
#| message: false

# Create analysis with coded negative results for Article 5 Sections 5.01-5.09
article5_likely_results <- data.frame(
  Category = c(
    "Charter School PEIMS Compliance Requirements",
    "Charter School Reading and Mathematics Requirements",
    "Charter School Accelerated Instruction Requirements",
    "Charter School Special Education Program Requirements",
    "Charter School Bilingual Education Requirements",
    "Charter School Prekindergarten Program Requirements",
    "Charter School Discipline Management Requirements",
    "Charter School Health and Safety Requirements",
    "Charter School Accountability Requirements",
    "Charter School Bullying Prevention Requirements",
    "Charter School Safety Requirements",
    "Charter School Early Childhood Literacy Plans",
    "Charter School College Career Military Readiness Plans",
    "Teacher Literacy Achievement Academy Requirements",
    "Teacher Literacy Achievement Academy Training Content",
    "Teacher Literacy Achievement Academy Stipend Eligibility",
    "Teacher Literacy Achievement Academy Evaluation Requirements",
    "Teacher Reading Intervention Academy Requirements",
    "Teacher Mathematics Achievement Academy Requirements",
    "Teacher Mathematics Achievement Academy Stipend Eligibility",
    "Teacher Mathematics Achievement Academy Evaluation Requirements",
    "Teacher Mathematics Interventionist Academy Requirements",
    "Additional Days School Year Planning Grant Program",
    "Extended Year Program Attendance Requirements",
    "Reading Intervention Program Attendance Requirements",
    "Kindergarten Reading Readiness Assessment Requirements",
    "Reading Instrument Adoption Requirements",
    "Reading Instrument Administration Requirements",
    "Kindergarten Readiness Performance Standards",
    "Reading Assessment Reporting Requirements",
    "School Readiness Certification System Requirements",
    "Early Literacy and Numeracy Instrument Requirements",
    "Reading Instrument Cost Requirements"
  ),

  Incentive_Created = c(
    "Open-enrollment charter schools subject to Public Education Information Management System requirements to extent necessary to monitor compliance",
    "Charter schools must comply with reading and mathematics instruments and reading interventions under specified sections",
    "Charter schools must provide accelerated instruction under state requirements for students needing additional support",
    "Charter schools must provide special education programs under state requirements equivalent to traditional public schools",
    "Charter schools must provide bilingual education under state requirements for English language learners",
    "Charter schools must provide prekindergarten programs under state requirements except class size limits do not apply",
    "Charter schools must implement discipline management practices and behavior management techniques under state oversight",
    "Charter schools must comply with health and safety requirements under state oversight and monitoring",
    "Charter schools subject to public school accountability and special investigations under multiple accountability subchapters",
    "Charter schools must implement bullying prevention policies and procedures and report incidents under state requirements",
    "Charter schools must comply with extensive school safety requirements under multiple sections and subchapters",
    "Charter schools must establish early childhood literacy and mathematics proficiency plans under state requirements",
    "Charter schools must establish college career and military readiness plans under state oversight and monitoring",
    "Teachers who provide reading instruction must include training in phonemic awareness phonics fluency vocabulary and use empirically validated methods",
    "Reading instruction training must include effective instructional practices for struggling readers and may include writing instruction training",
    "Classroom teachers who complete literacy achievement academy entitled to receive stipend from school district except under specified exemptions",
    "Agency must develop method for evaluating literacy achievement academy effectiveness including impact on teaching practices and student proficiency",
    "Commissioner must develop and make available reading intervention academies for teachers providing reading interventions to students requiring foundational skills",
    "Teachers who provide mathematics instruction must complete mathematics achievement academy and receive training in effective instructional practices",
    "Classroom teachers who complete mathematics achievement academy entitled to receive stipend from school district except under specified exemptions",
    "Agency must develop method for evaluating mathematics achievement academy effectiveness including impact on teaching practices and student proficiency",
    "Commissioner must develop and make available mathematics interventionist academies for teachers providing mathematics interventions requiring foundational skills",
    "Agency must establish and administer grant program to provide funding and technical assistance for school year planning to qualify for incentive funding",
    "Students must attend extended-year programs tutorial classes reading intervention programs accelerated instruction programs or basic skills programs",
    "Students assigned to reading intervention programs must attend unless specifically exempted under state requirements",
    "Commissioner must develop recommendations for school districts for administering reading instruments and training educators in administration",
    "Commissioner must adopt reading instrument that school districts must use to measure kindergarten foundational literacy skills in reading development",
    "School districts must administer reading instrument at kindergarten level not later than 60th day after beginning of school year",
    "Commissioner by rule must determine performance on reading instrument that indicates kindergarten readiness and provide comparison ability",
    "Superintendent must report reading instrument results to commissioner and board at public meeting and to parents within specified timeframes",
    "Districts must report electronically each student raw score on reading instrument to agency for school readiness certification system",
    "Commissioner must adopt list of reading and mathematics instruments for measuring foundational literacy skills in reading development and comprehension",
    "Agency must ensure reading instruments adopted or approved under section are available to school districts at no cost"
  ),

  Intended_Consequence = c(
    "Ensure charter schools provide data necessary for state monitoring while maintaining operational flexibility",
    "Ensure charter schools provide equivalent reading and mathematics instruction quality as traditional public schools",
    "Ensure charter school students receive appropriate accelerated instruction when needed for academic success",
    "Ensure charter school students with disabilities receive appropriate special education services equivalent to traditional schools",
    "Ensure charter school English language learners receive appropriate bilingual education services and support",
    "Provide charter schools flexibility in prekindergarten class sizes while maintaining program quality standards",
    "Ensure charter schools maintain appropriate student discipline and behavior management aligned with state standards",
    "Ensure charter schools maintain safe and healthy learning environments equivalent to traditional public schools",
    "Hold charter schools accountable for student performance and educational quality through comprehensive accountability systems",
    "Ensure charter schools provide safe learning environments free from bullying through prevention and reporting requirements",
    "Ensure charter schools maintain comprehensive safety protocols and emergency preparedness equivalent to traditional schools",
    "Ensure charter schools develop systematic approaches to early childhood literacy and mathematics instruction",
    "Ensure charter schools prepare students for post-secondary success through college career and military readiness planning",
    "Improve early elementary reading instruction quality through systematic teacher training in evidence-based practices",
    "Enhance teacher capability to support struggling readers through specialized training in empirically validated methods",
    "Provide financial incentive for teachers to complete professional development while ensuring district discretion over participation",
    "Ensure literacy achievement academies actually improve teaching practices and student outcomes through systematic evaluation",
    "Provide specialized training for teachers working with students requiring intensive reading intervention support",
    "Improve mathematics instruction quality through systematic teacher training in effective instructional practices",
    "Provide financial incentive for teachers to complete mathematics professional development while ensuring district discretion",
    "Ensure mathematics achievement academies actually improve teaching practices and student outcomes through systematic evaluation",
    "Provide specialized training for teachers working with students requiring intensive mathematics intervention support",
    "Support districts in planning extended school year operations to maximize state incentive funding opportunities",
    "Ensure students receive intensive academic support through mandatory participation in remedial and intervention programs",
    "Ensure students identified as needing reading intervention receive appropriate intensive instruction to improve literacy",
    "Provide systematic guidance to districts for implementing effective reading assessment and instruction practices",
    "Establish consistent statewide approach to measuring kindergarten reading readiness and foundational literacy skills",
    "Ensure timely assessment of kindergarten students to identify those needing early literacy intervention support",
    "Create standardized definition of kindergarten readiness to guide instructional decisions and resource allocation",
    "Ensure transparency in reading assessment results and timely communication with parents about student progress",
    "Integrate individual student reading data into statewide certification system for comprehensive tracking",
    "Provide systematic approach to measuring foundational literacy and numeracy skills across early elementary grades",
    "Remove financial barriers for districts to access high-quality reading assessment instruments approved by state"
  ),

  Likely_Result = c(
    "C, G, I, L, M", # Charter PEIMS compliance: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Charter reading/math requirements: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter accelerated instruction: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter special education: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter bilingual education: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter prekindergarten: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter discipline management: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter health and safety: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Charter accountability: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Charter bullying prevention: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Charter safety requirements: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Charter literacy plans: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Charter readiness plans: failure by force, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Teacher literacy academy requirements: failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Teacher literacy training content: failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "A, B, C, F, G, I, L", # Teacher literacy stipend: intrinsic→extrinsic, money focus, failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Teacher literacy evaluation: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, F, G, I, L", # Reading intervention academies: failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "C, F, G, I, L", # Teacher math academy requirements: failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "A, B, C, F, G, I, L", # Teacher math stipend: intrinsic→extrinsic, money focus, failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Teacher math evaluation: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, F, G, I, L", # Math interventionist academies: failure by force, reduce system improvement focus, reduced local control, system tampering, bureaucracy
    "A, B, C, G, I, K, L, M", # School year planning grants: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, tax burden, bureaucracy, waste resources
    "C, G, I, L", # Extended year attendance: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Reading intervention attendance: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Reading readiness recommendations: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L", # Reading instrument adoption: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Reading instrument administration: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Kindergarten readiness standards: failure by force, reduced local control, system tampering, bureaucracy
    "C, G, I, L, M", # Reading assessment reporting: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L, M", # School readiness certification: failure by force, reduced local control, system tampering, bureaucracy, waste resources
    "C, G, I, L", # Early literacy instruments: failure by force, reduced local control, system tampering, bureaucracy
    "A, B, C, G, I, K, L, M" # Reading instrument cost: intrinsic→extrinsic, money focus, failure by force, reduced local control, system tampering, tax burden, bureaucracy, waste resources
  )
)

# Verify all vectors have same length
stopifnot(length(article5_likely_results$Category) == length(article5_likely_results$Incentive_Created))
stopifnot(length(article5_likely_results$Incentive_Created) == length(article5_likely_results$Intended_Consequence))
stopifnot(length(article5_likely_results$Intended_Consequence) == length(article5_likely_results$Likely_Result))

# Create the main analysis table with coded results
article5_main_table <- datatable(
  article5_likely_results,
  options = list(
    pageLength = 10,
    scrollX = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '150px', targets = 0),   # Category  
      list(width = '300px', targets = 1),   # Incentive Created
      list(width = '250px', targets = 2),   # Intended Result
      list(width = '150px', targets = 3)    # Coded Results
    ),
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel')
  ),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover',
  colnames = c('Category', 'Incentive Created', 'Intended Result', 'Codes for Destructive Results We Will Actually Get')
) %>%
  formatStyle(
    'Category',
    backgroundColor = styleEqual(
      c("Charter School PEIMS Compliance Requirements", "Charter School Reading and Mathematics Requirements", "Charter School Accelerated Instruction Requirements", 
        "Charter School Special Education Program Requirements", "Charter School Bilingual Education Requirements", "Charter School Prekindergarten Program Requirements",
        "Charter School Discipline Management Requirements", "Charter School Health and Safety Requirements", "Charter School Accountability Requirements",
        "Charter School Bullying Prevention Requirements", "Charter School Safety Requirements", "Charter School Early Childhood Literacy Plans",
        "Charter School College Career Military Readiness Plans", "Teacher Literacy Achievement Academy Requirements", "Teacher Literacy Achievement Academy Training Content",
        "Teacher Literacy Achievement Academy Stipend Eligibility", "Teacher Literacy Achievement Academy Evaluation Requirements", "Teacher Reading Intervention Academy Requirements",
        "Teacher Mathematics Achievement Academy Requirements", "Teacher Mathematics Achievement Academy Stipend Eligibility", "Teacher Mathematics Achievement Academy Evaluation Requirements",
        "Teacher Mathematics Interventionist Academy Requirements", "Additional Days School Year Planning Grant Program", "Extended Year Program Attendance Requirements",
        "Reading Intervention Program Attendance Requirements", "Kindergarten Reading Readiness Assessment Requirements", "Reading Instrument Adoption Requirements",
        "Reading Instrument Administration Requirements", "Kindergarten Readiness Performance Standards", "Reading Assessment Reporting Requirements",
        "School Readiness Certification System Requirements", "Early Literacy and Numeracy Instrument Requirements", "Reading Instrument Cost Requirements"),
      c('#FFE8E8', '#FFD0D0', '#E8F4FD', '#E8F6FF', '#FFF2CC', '#E8F5E8', '#F0E8FF', '#FFE8F5', 
        '#FFEBCD', '#F0FFF0', '#F5F5DC', '#FFFACD', '#F0F8FF', '#FDF5E6', '#FAF0E6', '#FFF8DC', 
        '#F5FFFA', '#FFEFD5', '#E6E6FA', '#F0FFFF', '#FFF0F5', '#F5F5F5', '#FFFAF0', '#F8F8FF',
        '#FFFFF0', '#FDF5E6', '#FAF0E6', '#FFF8DC', '#F5FFFA', '#FFEFD5', '#E6E6FA', '#F0FFFF', '#FFF0F5')
    ),
    fontWeight = 'bold'
  ) %>%
  formatStyle(
    'Likely_Result',
    fontWeight = 'bold',
    fontSize = '14px',
    textAlign = 'center',
    backgroundColor = '#FFF8DC'
  )

# Display article 5 main table
article5_main_table
```

<br>
<br>
<br>

### Supplemental Tables

```{r}
#| label: supplemental-analysis-article5-sections01-09
#| echo: false
#| warning: false
#| message: false

# Create category analysis
category_analysis_a5 <- article5_likely_results %>%
  group_by(Category) %>%
  summarise(
    Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(desc(Count)) %>%
  rename(`Incentive Category` = Category,
         `Number of Incentives` = Count)

# Analyze most frequent negative result codes
code_frequency_a5 <- article5_likely_results %>%
  separate_rows(Likely_Result, sep = ", ") %>%
  count(Likely_Result, sort = TRUE) %>%
  rename(`Negative Result Code` = Likely_Result,
         `Frequency` = n) %>%
  mutate(Percentage = round(100 * Frequency / nrow(article5_likely_results), 3))

# Key policy findings focused on coded patterns
key_findings_a5 <- data.frame(
  Finding = c(
    "Most Common Negative Pattern",
    "System Tampering Universal Impact", 
    "Bureaucratic Administrative Burden",
    "Financial Motivation Corruption Pattern",
    "Resource Waste in Assessment Systems",
    "Most Problematic Policy Areas",
    "State Centralization Complete Takeover", 
    "Charter School Regulatory Burden Pattern",
    "Teacher Professional Development Corruption"
  ),
  Description = c(
    "Code C (failure by force) appears in 100.0% of incentives, and Code G (reduced local control) appears in 100.0%, indicating universal coercion and centralization",
    "Code I (system tampering) appears in 100.0% of incentives, showing complete disruption of existing educational systems for early childhood education",
    "Code L (bureaucratic burdens) appears in 100.0% of incentives, creating massive compliance overhead across all early childhood education policies",
    "Code A (intrinsic to extrinsic motivation) appears in 12.1% of incentives, systematically undermining professional dedication through financial incentives in teacher development",
    "Code M (waste resources) appears in 30.3% of incentives, concentrated in assessment systems, reporting requirements, and program monitoring",
    "Charter School compliance requirements and Teacher Achievement Academy policies show highest concentrations with 13 and 11 harmful incentives respectively",
    "Code G (reduced local control) appears in 100.0% of incentives, representing complete state takeover of early childhood education decisions", 
    "Charter schools subject to identical state control as traditional public schools while losing operational autonomy that was their primary advantage",
    "Code F (reduce system improvement focus) appears in 27.3% of incentives, making individual teacher training the solution rather than improving working conditions"
  )
)

# === DISPLAY SUPPLEMENTAL TABLES ===

# Category Analysis Header
datatable(
  data.frame(Title = "INCENTIVE CATEGORY ANALYSIS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#FF9800',
              color = 'white')

# Category Analysis Table
datatable(
  category_analysis_a5,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Number of Incentives',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(category_analysis_a5$`Number of Incentives`), '#E8F4FD')) %>%
  formatStyle('Incentive Category',
              fontWeight = 'bold',
              textAlign = 'left')

# Code Frequency Analysis Header
datatable(
  data.frame(Title = "NEGATIVE RESULT CODE FREQUENCY"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#DC143C',
              color = 'white')

# Code Frequency Table
datatable(
  code_frequency_a5,
  filter = 'top',
  options = list(
    pageLength = 15,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Frequency',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = styleColorBar(range(code_frequency_a5$Frequency), '#FFB6C1')) %>%
  formatStyle('Negative Result Code',
              fontWeight = 'bold',
              textAlign = 'center',
              fontSize = '16px') %>%
  formatStyle('Percentage',
              fontWeight = 'bold',
              textAlign = 'center')

# Key Findings Header
datatable(
  data.frame(Title = "KEY CODED PATTERN FINDINGS"),
  filter = 'top',
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#F44336',
              color = 'white')

# Key Findings Table
datatable(
  key_findings_a5,
  filter = 'top',
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '25%', targets = 0),
      list(width = '75%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Finding',
              fontWeight = 'bold',
              backgroundColor = '#FFE4E1',
              textAlign = 'left') %>%
  formatStyle('Description',
              textAlign = 'left')

# Create coded summary statistics
summary_statistics_a5 <- data.frame(
  Metric = c(
    "Total Incentives Analyzed",
    "Most Problematic Code", 
    "Most Frequent Code Combination",
    "Policy Categories with Code C",
    "Policy Categories with Code G",
    "Average Codes per Incentive"
  ),
  Value = c(
    nrow(article5_likely_results),
    paste0("Code C, G, I, L (all appear ", max(code_frequency_a5$Frequency), " times each)"),
    "C, G, I, L (force + control + tampering + bureaucracy)",
    sum(grepl("C", article5_likely_results$Likely_Result)),
    sum(grepl("G", article5_likely_results$Likely_Result)),
    round(mean(lengths(strsplit(article5_likely_results$Likely_Result, ", "))), 1)
  )
)

# Summary Statistics Header
datatable(
  data.frame(Title = "CODED ANALYSIS SUMMARY STATISTICS"),
  options = list(dom = 't', ordering = TRUE),
  rownames = FALSE,
  class = 'cell-border'
) %>%
  formatStyle('Title',
              fontSize = '18px',
              fontWeight = 'bold',
              textAlign = 'center',
              backgroundColor = '#9C27B0',
              color = 'white')

# Summary Statistics Table
datatable(
  summary_statistics_a5,
  options = list(
    pageLength = 10,
    dom = 't',
    ordering = TRUE,
    autoWidth = TRUE,
    columnDefs = list(
      list(width = '40%', targets = 0),
      list(width = '60%', targets = 1)
    )
  ),
  rownames = FALSE,
  class = 'cell-border stripe'
) %>%
  formatStyle('Metric',
              fontWeight = 'bold',
              textAlign = 'left') %>%
  formatStyle('Value',
              textAlign = 'left',
              fontWeight = 'bold')
```

<br>
<br>

### Key to List of Likely Negative Results

```{r}
#| label: negative-results-coding-key-article5-sections01-09
#| echo: false
#| message: false
#| warning: false

# Create negative results table with color coding
negative_results_a5 <- data.frame(
  Code = c("A", "B", "C", "D", "E", "F", "G", "H", "I", "J", "K", "L", "M", "N"),

  Likely_Negative_Result = c(
    "Will move people from intrinsic to extrinsic motivation",
    "Will makes money the most important motivation",  
    "Will fail because improvements cannot be realized by force from the state",
    "People will become 'the problem', not the system that they work in",
    "Will decrease collaboration, pitting people against each other for rewards",
    "Will decrease opportunities and efforts to improve the system teachers work in; try to 'improve' individual teachers instead",
    "Will reduce local control",
    "Will reduce economic agility of districts when they need it most",
    "Will tamper with existing stable systems, making things worse",
    "Impedes career development when goal posts constantly moving",
    "Shifts the onus of tax increases to districts to pay for unfunded state mandates",
    "Will create more administrative and bureaucratic burdens",
    "Will cause administration and teachers to waste even more resources, while bankrupting financially responsible districts",
    "Will create long-term district obligations that are not sustainable when state funding is changed every 2 years"
  )
)

# Define bright contrasting color palette (7 colors)
color_palette_a5 <- c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF")

# Create color assignments for each row (cycling through 7 colors for 14 rows)
row_colors_a5 <- rep(color_palette_a5, length.out = nrow(negative_results_a5))

# Display the coding key table
DT::datatable(
  negative_results_a5,
  options = list(
    pageLength = 15,
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel'),
    columnDefs = list(
      list(width = '40px', targets = 0),  # Narrow first column
      list(width = '90%', targets = 1)    # Wide second column
    ),
    scrollX = FALSE,
    autoWidth = FALSE
  ),
  rownames = FALSE,
  class = 'display'
) %>%
  formatStyle(
    'Code',
    textAlign = 'right',
    fontWeight = 'bold',
    backgroundColor = styleEqual(negative_results_a5$Code, row_colors_a5),
    color = 'white'
  ) %>%
  formatStyle(
    'Likely_Negative_Result',
    backgroundColor = styleEqual(negative_results_a5$Code, row_colors_a5),
    color = 'black'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    border = '2px solid #333',
    borderCollapse = 'collapse'
  ) %>%
  formatStyle(
    columns = c(0, 1),
    maxWidth = '9in',
    tableLayout = 'fixed'
  )

```

